---
title: "1 Creation of GNI Database"
author: "Martina Pesce & Edward P. K. Parker"
date: "2023-10-02"
output: html_document
editor_options: 
  chunk_output_type: console
---


#### Inputs
- Input/inequalities databases/GNI per capita (ppp)/GNI_per_c.csv: country-level GNI per capita
- Input/inequalities databases/income_group (World Bank).xlsx: World Bank regions and income groups

#### Outputs
- Processed/Ineq_merged.csv: Combined country-level income data

#### Notes
GNI per capita represents the primary covariate of interest. This indicator provides per capita values for gross national income (GNI. Formerly GNP) expressed in current international dollars converted by purchasing power parity (PPP) conversion factor (World Bank). 
World Bank income groups provide additional descriptive information.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
#libraries

library(readxl)
library(countrycode)
library(tidyverse)
```

# Inequalities Data

```{r import ineq database}
# Remove the first 4 lines of metadata
skip_lines <- 4

# Read the CSV file, skipping the metadata lines 

#GNI
GNI_per_c_WB <- read.csv("Data/Input/inequalities databases/WB_GNI_Atlas/API_NY.GNP.PCAP.CD_DS2_en_csv_v2_75655.csv", skip = skip_lines) %>%
  select(-c(X2020,X2021,X2022, X2023)) # remove years post COVID emergence

#World bank quartiles
Income_group_WB <- read_excel("Data/Input/inequalities databases/income_group (World Bank).xlsx")
```

```{r}
# Select latest GNI data 
GNI_per_c_WB_ok = GNI_per_c_WB %>%
  select(-`Country.Code`, -`Indicator.Name`, -`Indicator.Code`) %>% 
  # extract annual GNI data by country
  pivot_longer(cols = starts_with("X"), names_to = "year", values_to = "inequity_value", values_drop_na = TRUE) %>%
  mutate(year = as.integer(str_remove(year, "X")),
       inequity_value = as.numeric(inequity_value)) %>%
  # group data by country and select latest data
  group_by(`Country.Name`) %>%
  filter(year == max(year, na.rm = TRUE)) %>%
  ungroup() %>%
  select(`Country.Name`, inequity_value, year)

# rename final two columns with values from parameters 2/3
names(GNI_per_c_WB_ok)[2:3] = c("GNI_per_c_WB", "year_GNI")

# List countries without GNI data
GNI_per_c_WB$Country.Name[!(GNI_per_c_WB$Country.Name) %in% GNI_per_c_WB_ok$Country.Name] #n = 11

#  [1] "American Samoa"            "Channel Islands"           "Gibraltar"                 "Guam"                      "Not classified"      
#  [6] "St. Martin (French part)"  "Monaco"                    "Northern Mariana Islands"  "Korea, Dem. People's Rep." "British Virgin Islands" 
#  [11] "Virgin Islands (U.S.)"    

```

```{r FINAL inequalities}
# MERGE AND COUNTRY CODE
merge_ineq <- GNI_per_c_WB %>%
  select(Country.Name, Country.Code)

# Cross-check that country names in derived datasets are all represented in combined data
all(GNI_per_c_WB_ok$Country.Name %in% merge_ineq$Country.Name)

merge_ineq <- left_join(merge_ineq, GNI_per_c_WB_ok, by = "Country.Name")%>%
  mutate(iso3c = countrycode(Country.Name, origin = "country.name", destination = "iso3c"))

merge_ineq <- left_join(merge_ineq, Income_group_WB, by=c("Country.Code" = "Code")) %>%
  select(-`Lending category`)

# Summarise and filter
nrow(merge_ineq) # 266 countries or regions in World Bank data

# Filter regions (e.g. 'Africa Easter and Southern', 'Arab World')
merge_ineq <- merge_ineq %>% filter(!is.na(Economy))
nrow(merge_ineq) # 217 countries

# Filter countries without GNI data
merge_ineq <- merge_ineq %>% filter(!is.na(GNI_per_c_WB))
nrow(merge_ineq) # 207 countries

#Create quartiles
merge_ineq <- merge_ineq %>%
  # excluded Marshall Islands, the Federated States of Micronesia, and Puerto Rico from descriptive analyses as vaccine coverage metrics were not included in WHO or Our World in Data databases despite records of vaccine introductions in the public domain
  filter(Country.Name != "Micronesia, Fed. Sts." & Country.Name != "Marshall Islands" & Country.Name != "Puerto Rico") %>% 
  mutate(
    q_GNI_per_c_WB = ntile(GNI_per_c_WB, 4)
  )   

# Check that all countries have quartile assigned
sum(is.na(merge_ineq$q_GNI_per_c_WB)) # 0
```

```{r}
# Summarise year of GNI data
table(merge_ineq$year_GNI) # data pre 2019 for 7/204 (3%)

# Show countries with pre-2019 data
subset(merge_ineq, year_GNI!=2019)[,c("Country.Name", "year_GNI")]
```

```{r}
#Table q_GNI_per_c_WB vs. income group (WB)
table <- merge_ineq %>% 
  group_by(q_GNI_per_c_WB, `Income group`) %>% 
  summarise(count=n()) %>% 
  pivot_wider(
    names_from = `Income group`,
    values_from = count
  )

#GNI summary by quartile
summary_GNI <- merge_ineq %>%
  group_by(q_GNI_per_c_WB) %>%
  summarize(
    n = length(q_GNI_per_c_WB),
    median_value = round(median(GNI_per_c_WB, na.rm=TRUE),0),
    q25 = round(quantile(GNI_per_c_WB, 0.25, na.rm=TRUE),0),
    q75 = round(quantile(GNI_per_c_WB, 0.75, na.rm=TRUE),0)
  )
```


```{r, write inequalities data}
#Save inequalities data

write.csv(merge_ineq, "Data/Processed/Ineq_merged.csv")


```

# Session info
```{r, session info}
print(sessionInfo())
```

