---
title: "2 Introduction and scale-up description (Fig 1)"
author: "Martina Pesce & Edward P. K. Parker"
date: "2023-11-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

#### Inputs
- Processed/Ineq_merged.csv: Combined country-level income data
- Input/Vaccine_data/owid-covid-data.csv: Our World in Data COVID-19 database
- Input/Vaccine_data/owid_vaccine_plat.csv: Our World in Data COVID-19 vaccine platforms
- Input/Vaccine_data/vaccination-data.csv: WHO COVID-19 vaccine database

#### Outputs
- Processed/Ineq_merged_with_vaccine_data.csv: Country-level income data with vaccine introduction and scale-up metrics 
- Descriptive statistics on vaccine introduction and scale-up
- Figure 1


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r Libraries}
library(tidyverse)
library(readr)
library(sf)
library(rnaturalearth)
library(RColorBrewer)
library(lubridate)
library(countrycode)
library(patchwork)
 
```



```{r Data base F1}

#Ineq database GNI
Ineq_merged <- read_csv("Data/Processed/Ineq_merged.csv") %>% 
  select(Country.Name:q_GNI_per_c_WB)


#Coverage information

##Our world in data 
owid_covid_data <- read_csv("Data/Input/Vaccine_data/owid-covid-data_2024-01-07.csv") %>% 
select(
    iso_code:date, total_vaccinations_per_hundred: people_fully_vaccinated_per_hundred)

##WHO
WHO_covid_data <- read_csv("Data/Input/Vaccine_data/WHO_vaccination-data-2024-01-07.csv") %>%   
  select(
    ISO3, FIRST_VACCINE_DATE
  )

##WHO health care workers and older populations
WHO_hcw_old_data <- read_csv("Data/Input/Vaccine_data/WHO_UNICEF_data/who_unicef_country_timeseries_data.csv") %>% 
  select(name_short, iso, date, cov_hcw_cps, cov_old_cps, cov_tot_cps, cov_tot_boost, cov_hcw_boost, cov_old_boost,pop, pop_hcw, pop_old) %>% 
  mutate(date = ifelse(is.na(date), NA, paste0(date, "-01")))

WHO_hcw_old_sum_data <- read_csv("Data/Input/Vaccine_data/WHO_UNICEF_data/who_unicef_country_summary_data.csv") %>% 
  select(iso, pol_age_old, pol_age_old_source)



#vaccines platforms
owid_vaccine_plat <- read_csv("Data/Input/Vaccine_data/owid_vaccine_plat.csv") %>%
  select(location:source_website)

```


#Figure 1.

## Maps


```{r}
coverage <- owid_covid_data %>%
  group_by(location) %>%
  mutate(
    owid_vac_intro = ifelse(row_number() == which(total_vaccinations_per_hundred > 0)[1], as.character(date), NA),
    owid_vac_F40 = ifelse(row_number() == which(people_fully_vaccinated_per_hundred > 39.99)[1], as.character(date), NA),
    owid_vac_F70 = ifelse(row_number() == which(people_fully_vaccinated_per_hundred > 69.99)[1], as.character(date), NA)
  ) %>%
  filter(!is.na(owid_vac_intro) | !is.na(owid_vac_F40) | !is.na(owid_vac_F70)) %>% 
  select(!(date:people_fully_vaccinated_per_hundred)) %>% 
  fill(owid_vac_intro:owid_vac_F70)%>%
  slice_tail() %>%  
  ungroup()

# Merge data frames and calculate minimum vaccine campaign date (OWID vs WHO)
Ineq_vaccine_coverage <- Ineq_merged %>%
  left_join(coverage, by = c("Country.Code" = "iso_code")) %>%
  left_join(WHO_covid_data, by = c("Country.Code" = "ISO3")) %>%
  # Convert date columns to Date format
  mutate(
    who_vac_intro = as.Date(as.character(FIRST_VACCINE_DATE)),
    owid_vac_intro = as.Date(owid_vac_intro)
  ) %>%
  # Calculate the date difference and minimum date
  mutate(
    min_vac_intro = pmin(owid_vac_intro, who_vac_intro, na.rm = TRUE),
    diff_owid_who_intro = as.numeric(owid_vac_intro - who_vac_intro)
  )%>% 
    mutate(
    diff_min_vac_intro_8Dec2020 = as.numeric(min_vac_intro - dmy("8-12-2020")), 
    diff_intro_F40 = as.numeric(as.Date(owid_vac_F40) - as.Date(min_vac_intro)),
    diff_intro_F70 = as.numeric(as.Date(owid_vac_F70) - as.Date(min_vac_intro)),
    owid_who= case_when(
      diff_owid_who_intro < 0 | (is.na(who_vac_intro) & !is.na(owid_vac_intro))~ "owid",
      diff_owid_who_intro > 0 | (is.na(owid_vac_intro) & !is.na(who_vac_intro)) ~ "who",
      diff_owid_who_intro == 0 ~ "equal",
      TRUE ~ "no date"
    )
    ) %>% 
  mutate(
    diff_min_vac_intro_8Dec2020_clean=case_when(
      diff_min_vac_intro_8Dec2020<0 ~ 0,
      TRUE ~ diff_min_vac_intro_8Dec2020
    )
  ) %>%
  select(-FIRST_VACCINE_DATE, -location)

table(Ineq_vaccine_coverage$owid_who)

#median dates to built groups
summary_vacc_intro_scale <- Ineq_vaccine_coverage %>%
  
  group_by(q_GNI_per_c_WB) %>%
  summarize(
    # Count of countries in each GNI quartile
    n = length(q_GNI_per_c_WB),
    # Count of countries with available vaccine introduction dates
    n_intro = sum(!is.na(min_vac_intro)),
    # Summary statistics for introduction delays (in days since December 8, 2020)
    median_value_1 = round(median(diff_min_vac_intro_8Dec2020, na.rm = TRUE), 0),
    q25_1 = round(quantile(diff_min_vac_intro_8Dec2020, 0.25, na.rm = TRUE), 0),
    q75_1 = round(quantile(diff_min_vac_intro_8Dec2020, 0.75, na.rm = TRUE), 0),
    # Summary statistics for scale-up delays to reach 40% coverage (in days)
    median_value_40 = round(median(diff_intro_F40, na.rm = TRUE), 0),
    q25_40 = round(quantile(diff_intro_F40, 0.25, na.rm = TRUE), 0),
    q75_40 = round(quantile(diff_intro_F40, 0.75, na.rm = TRUE), 0),
    # Summary statistics for scale-up delays to reach 70% coverage (in days)
    median_value_70 = round(median(diff_intro_F70, na.rm = TRUE), 0),
    q25_70 = round(quantile(diff_intro_F70, 0.25, na.rm = TRUE), 0),
    q75_70 = round(quantile(diff_intro_F70, 0.75, na.rm = TRUE), 0),
    # Dates for vaccine introduction and scale-ups (median and quartiles)
    median_value_1_date = as.Date(median(as.numeric(min_vac_intro), na.rm = TRUE), origin = "1970-01-01"),
    q25_1_date = as.Date(quantile(as.numeric(min_vac_intro), 0.25, na.rm = TRUE), origin = "1970-01-01"),
    q75_1_date = as.Date(quantile(as.numeric(min_vac_intro), 0.75, na.rm = TRUE), origin = "1970-01-01"),
    # Dates for reaching 40% coverage
    median_value_40_date = as.Date(median(as.numeric(as.Date(owid_vac_F40)), na.rm = TRUE), origin = "1970-01-01"),
    q25_40_date = as.Date(quantile(as.numeric(as.Date(owid_vac_F40)), 0.25, na.rm = TRUE), origin = "1970-01-01"),
    q75_40_date = as.Date(quantile(as.numeric(as.Date(owid_vac_F40)), 0.75, na.rm = TRUE), origin = "1970-01-01"),
    # Dates for reaching 70% coverage
    median_value_70_date = as.Date(median(as.numeric(as.Date(owid_vac_F70)), na.rm = TRUE), origin = "1970-01-01"),
    q25_70_date = as.Date(quantile(as.numeric(as.Date(owid_vac_F70)), 0.25, na.rm = TRUE), origin = "1970-01-01"),
    q75_70_date = as.Date(quantile(as.numeric(as.Date(owid_vac_F70)), 0.75, na.rm = TRUE), origin = "1970-01-01")
  ) %>%
  mutate(
    # Create readable summary columns for delays
    introduction = paste0(median_value_1, " (", q25_1, "-", q75_1, ")"),
    scale_40 = paste0(median_value_40, " (", q25_40, "-", q75_40, ")"),
    scale_70 = paste0(median_value_70, " (", q25_70, "-", q75_70, ")"),
    # Create readable summary columns for dates
    introduction_date = paste0(format(median_value_1_date, "%d/%m/%Y"), " (", 
                                format(q25_1_date, "%d/%m/%Y"), "-", 
                                format(q75_1_date, "%d/%m/%Y"), ")"),
    scale_40_date = paste0(format(median_value_40_date, "%d/%m/%Y"), " (", 
                           format(q25_40_date, "%d/%m/%Y"), "-", 
                           format(q75_40_date, "%d/%m/%Y"), ")"),
    scale_70_date = paste0(format(median_value_70_date, "%d/%m/%Y"), " (", 
                           format(q25_70_date, "%d/%m/%Y"), "-", 
                           format(q75_70_date, "%d/%m/%Y"), ")")
  ) %>%
  # Select and organize the relevant output columns
  select(
    q_GNI_per_c_WB, n, n_intro, 
    introduction, scale_40, scale_70, 
    introduction_date, scale_40_date, scale_70_date
  )

summary_vacc_intro_scale


#First vaccine introduction after 08-12-2020 by quartile

min_intro <- Ineq_vaccine_coverage %>%
  group_by(q_GNI_per_c_WB) %>%
  summarize(
min_vac_intro=min(as.Date(min_vac_intro)[!is.na(min_vac_intro) & as.Date(min_vac_intro) > as.Date("2020-12-07")])
  )

min_intro <- left_join(min_intro, Ineq_vaccine_coverage) %>% 
  select(1:3)

min_intro



#median dates to built groups
summary_vacc_dates <- Ineq_vaccine_coverage %>%
  # mutate(
  #   min_vac_intro_month = floor_date(as.Date(min_vac_intro), "month"),
  #   vac_F40_month = floor_date(as.Date(owid_vac_F40), "month"),
  #   vac_F70_month = floor_date(as.Date(owid_vac_F70), "month")
  # ) %>%
  group_by(q_GNI_per_c_WB) %>%
  summarize(
    n = length(q_GNI_per_c_WB),
    n_intro = sum(!is.na(min_vac_intro)),
    median_value_1 = as.Date(median(as.numeric(min_vac_intro), na.rm = TRUE), origin = "1970-01-01"),
    q25_1 = as.Date(quantile(as.numeric(min_vac_intro), 0.25, na.rm = TRUE), origin = "1970-01-01"),
    q75_1 = as.Date(quantile(as.numeric(min_vac_intro), 0.75, na.rm = TRUE), origin = "1970-01-01"),
    median_value_40 = as.Date(median(as.numeric(as.Date(owid_vac_F40)), na.rm = TRUE), origin = "1970-01-01"),
    q25_40 = as.Date(quantile(as.numeric(as.Date(owid_vac_F40)), 0.25, na.rm = TRUE), origin = "1970-01-01"),
    q75_40 = as.Date(quantile(as.numeric(as.Date(owid_vac_F40)), 0.75, na.rm = TRUE), origin = "1970-01-01"),
    median_value_70 = as.Date(median(as.numeric(as.Date(owid_vac_F70)), na.rm = TRUE), origin = "1970-01-01"),
    q25_70 = as.Date(quantile(as.numeric(as.Date(owid_vac_F70)), 0.25, na.rm = TRUE), origin = "1970-01-01"),
    q75_70 = as.Date(quantile(as.numeric(as.Date(owid_vac_F70)), 0.75, na.rm = TRUE), origin = "1970-01-01")
  ) %>% 
  mutate(
    introduction = paste0( median_value_1, " (", q25_1, "-", q75_1, ")"),
    scale_40 = paste0( median_value_40, " (", q25_40, "-", q75_40, ")"),
    scale_70 = paste0( median_value_70, " (", q25_70, "-", q75_70, ")")
  ) %>% 
  select(q_GNI_per_c_WB, introduction, scale_40, scale_70) 


```




```{r}
#Create groups 
Ineq_vaccine_coverage <- Ineq_vaccine_coverage %>% 
  mutate(
            V1_date_group=  case_when(
      min_vac_intro < dmy("01-01-2021")  ~ "before jan/2021",
      min_vac_intro >= dmy("01-01-2021") & min_vac_intro < dmy("01-02-2021") ~ "jan/2021",
      min_vac_intro >= dmy("01-02-2021") & min_vac_intro < dmy("01-03-2021") ~ "feb/2021",
      min_vac_intro >= dmy("01-03-2021") ~ "march/2021 +",
      TRUE ~ "not reached"
  ),
            V1_group= case_when(
      diff_min_vac_intro_8Dec2020 < 30  ~ "<30",
      diff_min_vac_intro_8Dec2020 > 29 & diff_min_vac_intro_8Dec2020 < 60 ~ "30-59",
      diff_min_vac_intro_8Dec2020 > 59 & diff_min_vac_intro_8Dec2020 < 90 ~ "60-89",
      diff_min_vac_intro_8Dec2020 > 89 ~ "90+",
      TRUE ~ "not reached"
  ),
            V40_date_group=  case_when(
      owid_vac_F40 < dmy("01-06-2021")  ~ "before Jun/2021",
      owid_vac_F40 >= dmy("01-06-2021") & owid_vac_F40 < dmy("01-09-2021") ~ "Jun-Aug/2021",
      owid_vac_F40 >= dmy("01-09-2021") & owid_vac_F40 < dmy("01-12-2021") ~ "Sep-Nov/2021",
      owid_vac_F40 >= dmy("01-12-2021") ~ "Dec/2021 +",
      TRUE ~ "not reached"
  ),
            V40_group= case_when(
      diff_intro_F40 < 180  ~ "<180",
      diff_intro_F40 > 179 & diff_intro_F40 < 270 ~ "180-269",
      diff_intro_F40 > 269 & diff_intro_F40 < 360 ~ "270-359",
      diff_intro_F40 > 359 ~ "360+",
      TRUE ~ "not reached"
  ),
    V1_dummy = case_when(
      !is.na(min_vac_intro) & min_vac_intro < as.Date("2022-01-01") ~ "yes",
      TRUE ~ "no"
    ),
    V40_dummy = case_when(
      !is.na(owid_vac_F40) & owid_vac_F40 < as.Date("2022-01-01") ~ "yes",
      TRUE ~ "no"
    ),
    V70_dummy = case_when(
      !is.na(owid_vac_F70) & owid_vac_F70 < as.Date("2022-07-01") ~ "yes",
      TRUE ~ "no"
    ),
    V1_dummy_no_date = case_when(
      !is.na(min_vac_intro) ~ "yes",
      TRUE ~ "no"
    ),
    V40_dummy_no_date = case_when(
      !is.na(owid_vac_F40) ~ "yes",
      TRUE ~ "no"
    ),
    V70_dummy_no_date = case_when(
      !is.na(owid_vac_F70) ~ "yes",
      TRUE ~ "no"
    )
  )
  

#Geographical information

countries <-  ne_countries(returnclass = "sf") %>% 
  mutate(iso3c = countrycode(admin, origin = "country.name", destination = "iso3c")) %>% 
  filter(iso3c!="ATA")
  

world_ineq_gral <- 
  merge(countries, Ineq_vaccine_coverage, by.x = "iso3c", by.y = "Country.Code", all = TRUE)

world_ineq_gral <- st_as_sf(world_ineq_gral)


```



```{r}
#Map A.i: vaccine introduction

# map1_a<- ggplot(data = world_ineq_gral) +
#   geom_sf(aes(fill = as.factor(V1_group)), position = "identity") + 
#   scale_fill_manual(values = c("#2e73b8","#75b2d8","#9dd5ea", "#d7e0f4","#DDDDDD"),
#                     labels = c("<30", "30-59", "60-89", "90+", "not reached"),
#                     breaks = c("<30", "30-59", "60-89", "90+", "not reached")) +
#   labs(
#     title= "(A)\nTime to vaccine introduction",
#     fill = "Time since\n08 Dec 2020\n    (days)") +
#   theme_void()+
#   guides(color = guide_legend(title.position = "bottom", nrow = 1))+
#   theme(
#     legend.position = "bottom",
#     legend.text = element_text(size = 9),  # Adjust text size
#     legend.title = element_text(size = 8),  # Adjust legend title size
#     plot.title = element_text(size = 10) 
#   )
# 
# map1_a


#Map A.i.b: vaccine introduction date!

map1_a<- ggplot(data = world_ineq_gral) +
  geom_sf(aes(fill = as.factor(V1_date_group)), position = "identity") + 
  scale_fill_manual(values = c("#2e73b8","#75b2d8","#9dd5ea", "#d7e0f4","#A2B1A4"),
                    labels = c("< jan\n2021", "jan\n2021", "feb\n2021", "march\n2021 +", "not\nreached"),
                    breaks = c("before jan/2021", "jan/2021", "feb/2021", "march/2021 +", "not reached"),
                    na.value = "#F0F3F0") +
  labs(
    title= "(A)\nVaccine introduction date",
    fill = "Introduction\nmonth") +
  theme_void()+
  guides(color = guide_legend(title.position = "bottom", nrow = 1))+
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9),  # Adjust text size
    legend.title = element_text(size = 8),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )

map1_a
# 
# #Map A.ii: scale-up 
# map1_b<- ggplot(data = world_ineq_gral) +
#   geom_sf(aes(fill = as.factor(V40_group)), position = "identity") + 
#   scale_fill_manual(values = c("#827717","#afb42b", "#cddc39","#ecebb1","#DDDDDD"),
#                     breaks = c("<180", "180-269", "270-359", "360+", "not reached"),
#                     labels = c("<180", "180-269", "270-359", "360+", "not reached")) +
#   labs(
#            title= "Time to vaccine scale-up (40%)",
#     fill = "Time since\nintroduction\n      (days)") +
#   theme_void()+ 
#   theme(
#     legend.position = "bottom",
#     legend.text = element_text(size = 9),  # Adjust text size
#     legend.title = element_text(size = 8),  # Adjust legend title size
#     plot.title = element_text(size = 10) # Adjust plot title size
#   ) 
# 
# map1_b
#Map A.ii: scale-up DATE
map1_b<- ggplot(data = world_ineq_gral) +
  geom_sf(aes(fill = as.factor(V40_date_group)), position = "identity") + 
  scale_fill_manual(values = c("#827717","#afb42b", "#cddc39","#ecebb1","#A2B1A4"),
                    breaks = c("before Jun/2021", "Jun-Aug/2021", "Sep-Nov/2021", "Dec/2021 +", "not reached"),
                    labels = c("<Jun\n2021", "Jun-Aug\n2021", "Sep-Nov\n2021", "Dec\n2021 +", "not reached"),
                    na.value = "#F0F3F0") +
  labs(
           title= "Time of vaccine scale-up (40%)",
    fill = "Months") +
  theme_void()+ 
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9),  # Adjust text size
    legend.title = element_text(size = 8),  # Adjust legend title size
    plot.title = element_text(size = 10) # Adjust plot title size
  ) 

map1_b
```




# C: barplot and boxplots

```{r}
#ii) Bar plot introduction/scale-up yes no per GNI per capita quartile

milestone_achieved <- Ineq_vaccine_coverage %>% 
  group_by(q_GNI_per_c_WB) %>% 
  summarise(
    V1 = sum(V1_dummy == "yes"),
    V40 = sum(V40_dummy == "yes"),
    V70 = sum(V70_dummy == "yes"),
    V1_nd = sum(V1_dummy_no_date == "yes"),
    V40_nd = sum(V40_dummy_no_date == "yes"),
    V70_nd = sum(V70_dummy_no_date == "yes"),
    total = n(),
    V1_per = round(sum(V1_dummy == "yes") / n() * 100, 1),
    V40_per = round(sum(V40_dummy == "yes") / V1 * 100, 1),
    V70_per = round(sum(V70_dummy == "yes") / V1 * 100, 1),
    V1_per_nd = round(sum(V1_dummy_no_date == "yes") / n() * 100, 1),
    V40_per_nd = round(sum(V40_dummy_no_date == "yes") / V1_nd * 100, 1),
    V70_per_nd = round(sum(V70_dummy_no_date == "yes") / V1_nd * 100, 1),
  )

milestone_achieved

milestone_achieved <- milestone_achieved %>% 
    pivot_longer(
    cols = c(V1_per, V40_per),
    names_to = "milestone",
    values_to = "Value"
  )

milestone_achieved

```

```{r}
c1<- ggplot(milestone_achieved, aes(x = as.factor(q_GNI_per_c_WB), y = Value, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  coord_flip()+
  facet_wrap(~ milestone, labeller = labeller(milestone = c("V1_per" = "Vaccine introduction (31/12/2021)", "V40_per" = "Vaccine scale-up (40%, 31/12/2021)"))) +
  labs(title = "(B)",
       x = "",
       y = "% countries achieving milestone") +
    guides(fill = "none") +  # Remove the fill legend
  theme_minimal() +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal() + 
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),# Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) # Adjust plot title size
  ) 

c1



```


```{r}
# #ii) Box plot introduction/scale-up yes no per GNI per capita quartile
# intro_scale_delay<- Ineq_vaccine_coverage %>%
#   select(Country.Name, q_GNI_per_c_WB, diff_min_vac_intro_8Dec2020, diff_intro_F40) %>%
#   filter(!is.na(diff_min_vac_intro_8Dec2020)) %>%
#   pivot_longer(
#     cols = c(diff_min_vac_intro_8Dec2020, diff_intro_F40),
#     names_to = "Diff_dates",
#     values_to = "Value"
#   ) 
# 
# # Set factor levels
# intro_scale_delay$Diff_dates = factor(intro_scale_delay$Diff_dates, levels = unique(intro_scale_delay$Diff_dates))
# 
# c2_plot_intro_scale_delay <- ggplot(intro_scale_delay) +
#   geom_boxplot(aes(x = Value, y = as.factor(q_GNI_per_c_WB)), outlier.shape = NA) +
#    geom_point(
#     aes(
#      x = Value, 
#      y = as.factor(q_GNI_per_c_WB),
#       #size = studies,
#       color = as.factor(q_GNI_per_c_WB)
#     ),
#     alpha = 0.5,
#     position = position_jitterdodge(dodge.width = 0.5, jitter.width = 1) 
#   )+
#   scale_x_continuous(limits = c(0, NA))+
#   scale_color_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
#   facet_wrap(~Diff_dates, scales = "free_x", labeller = labeller(Diff_dates = c("diff_min_vac_intro_8Dec2020" = "Time to vaccine introduction", "diff_intro_F40" = "Time to vaccine scale-up (40%)")))+
#   labs(title = "(C)",
#        x = "Number of days", 
#        y = "GNI quartile")  +
#   guides(color = "none") +
#   scale_y_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
#   theme_minimal() + 
#   theme(
#     legend.position = "bottom",
#     legend.text = element_text(size = 8),# Adjust text size
#     legend.title = element_text(size = 9),  # Adjust legend title size
#     plot.title = element_text(size = 10) # Adjust plot title size
#   ) 
#  
# c2_plot_intro_scale_delay 

  
```


```{r}
#Option B
#ii) Box plot introduction/scale-up yes no per GNI per capita quartile
intro_scale_delay<- Ineq_vaccine_coverage %>%
  mutate(
    owid_vac_F40 = as.Date(owid_vac_F40)  ) %>% 
  select(Country.Name, q_GNI_per_c_WB, min_vac_intro, owid_vac_F40) %>%
  filter(!is.na(min_vac_intro)) %>%
  pivot_longer(
    cols = c(min_vac_intro, owid_vac_F40),
    names_to = "Date",
    values_to = "Value"
  ) 

# Set factor levels
 intro_scale_delay$Date = factor(intro_scale_delay$Date, levels = unique(intro_scale_delay$Date))

c2_plot_intro_scale_delay <- ggplot(intro_scale_delay) +
  geom_boxplot(aes(x = Value, y = as.factor(q_GNI_per_c_WB)), outlier.shape = NA) +
   geom_point(
    aes(
     x = Value, 
     y = as.factor(q_GNI_per_c_WB),
      #size = studies,
      color = as.factor(q_GNI_per_c_WB)
    ),
    alpha = 0.5,
    position = position_jitterdodge(dodge.width = 0.5, jitter.width = 1) 
  )+
  scale_x_date(limits = c(as.Date("2020-12-01"), NA),
               date_labels = "%b/%Y") +
  scale_color_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  facet_wrap(~Date, scales = "free_x", labeller = labeller(Date = c("min_vac_intro" = "Date of vaccine introduction", "owid_vac_F40" = "Date of vaccine scale-up (40%)")), nrow=2)+
  labs(title = "(C)",
       x = "Date", 
       y = "GNI quartile")  +
  guides(color = "none") +
  scale_y_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal() + 
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),# Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) # Adjust plot title size
  ) 
 
c2_plot_intro_scale_delay 

  
```


#1.D vaccine profile

#Panel 1.B 
Countries vaccine distribution and amount of papers produced by type and income quartile. 

```{r owid vacc plat}
#Data wrangling

# List any countries not represented in OWID dataset
Ineq_vaccine_coverage[!(Ineq_vaccine_coverage$iso3c %in% owid_covid_data$iso_code),]
# Kosovo - GNI quartile 2 - is not represented in OWID database
# 197 other countries in Ineq_vaccine_coverage are included in owid_covid_data (although not all have recorded vaccinations)

# List vaccines 
unique(unlist(strsplit(paste(owid_vaccine_plat$vaccines, collapse = ', '), ", ")))

# Create col for each vacc (1/0)
Ineq_vaccine_coverage_platform <- Ineq_vaccine_coverage %>%
  left_join(owid_vaccine_plat, by=c("Country.Code" = "iso_code")) %>%
  select(-location) %>%
  mutate(
    `CIGB CIGB-66/Abdala` = ifelse(str_detect(vaccines, "Abdala"), 1, 0),
    `Cansino Ad5-nCoV` = ifelse(str_detect(vaccines, "CanSino"), 1, 0),
    `Biological E Corbevax` = ifelse(str_detect(vaccines, "Corbevax"), 1, 0),
    `Bharat Covaxin/BBV152` = ifelse(str_detect(vaccines, "Covaxin"), 1, 0),
    `Shifa Pharmed COVIran` = ifelse(str_detect(vaccines, "COVIran Barekat"), 1, 0),
    `Chumakov Center CoviVac` = ifelse(str_detect(vaccines, "KoviVac/Chumakov"), 1, 0),
    `Vector Institute EpiVacCorona` = ifelse(str_detect(vaccines, "EpiVacCorona"), 1, 0),
    `Institute of Medical Biology CAMS vaccine` = ifelse(str_detect(vaccines, "IMBCAMS"), 1, 0),
    `Janssen Ad26.COV2.S` = ifelse(str_detect(vaccines, "Johnson&Johnson"), 1, 0),
    `Shenzhen Kangtai KCONVAC` = ifelse(str_detect(vaccines, "KCONVAC"), 1, 0),
    `Medicago CoVLP` = ifelse(str_detect(vaccines, "Medicago"), 1, 0),
    `Medigen MVC-COV1901` = ifelse(str_detect(vaccines, "Medigen"), 1, 0),
    `Moderna mRNA-1273` = ifelse(str_detect(vaccines, "Moderna"), 1, 0),
    `Novavax NVX-CoV2373` = ifelse(str_detect(vaccines, "Novavax"), 1, 0),
    `Oxford/AstraZeneca ChAdOx1-S` = ifelse(str_detect(vaccines, "Oxford/AstraZeneca") | str_detect(vaccines, "Covishield"), 1, 0),
    `BioNTech/Pfizer BNT162b2` = ifelse(str_detect(vaccines, "Pfizer/BioNTech"), 1, 0),
    `RIBSP Kazakhstan QazCovid-in` = ifelse(str_detect(vaccines, "QazVac"), 1, 0),
    `Beijing/Sinopharm BBIBP-CorV` = ifelse(str_detect(vaccines, "Sinopharm/Beijing") | str_detect(vaccines, "Sinopharm/HayatVax") | str_detect(vaccines, "BBIBP-CorV"), 1, 0),
    `Wuhan/Sinopharm vaccine` = ifelse(str_detect(vaccines, "Sinopharm/Wuhan"), 1, 0),
    `Sinovac CoronaVac` = ifelse(str_detect(vaccines, "Sinovac"), 1, 0),
    `Instituto Finlay de Vacunas Soberana 02` = ifelse(str_detect(vaccines, "Soberana02"), 1, 0),
    `Instituto Finlay de Vacunas Soberana 01` = ifelse(str_detect(vaccines, "Soberana Plus"), 1, 0),
    `Gamaleya Gam-COVID-Vac/Sputnik V` = ifelse(str_detect(vaccines, "Sputnik"), 1, 0),
    `Erciyes TURKOVAC` = ifelse(str_detect(vaccines, "Turkovac"), 1, 0),
    `AZLB ZF2001` = ifelse(str_detect(vaccines, "RBD-Dimer") | str_detect(vaccines, "ZF2001"), 1, 0),
    `Sanofi/GSK CoV2 preS dTM` = ifelse(str_detect(vaccines, "Sanofi/GSK"), 1, 0),
    `Valneva VLA2001` = ifelse(str_detect(vaccines, "Valneva"), 1, 0)
  ) %>%
  # Create col for each vacc (1/0)
  mutate(
   mRNA = ifelse(`BioNTech/Pfizer BNT162b2`==1 | `Moderna mRNA-1273`==1,1,0),
    Vectored = ifelse(`Oxford/AstraZeneca ChAdOx1-S`==1 | `Janssen Ad26.COV2.S`==1 | `Gamaleya Gam-COVID-Vac/Sputnik V`==1 | `Cansino Ad5-nCoV`==1,1,0),
    Inactivated = ifelse(`Beijing/Sinopharm BBIBP-CorV`==1 | `Sinovac CoronaVac`==1 | `Bharat Covaxin/BBV152`==1 | `Wuhan/Sinopharm vaccine`==1 | `RIBSP Kazakhstan QazCovid-in`==1 | 
                           `Shifa Pharmed COVIran`==1 | `Erciyes TURKOVAC`==1 | `Shenzhen Kangtai KCONVAC`==1 | `Institute of Medical Biology CAMS vaccine`==1 | `Chumakov Center CoviVac`==1 | `Valneva VLA2001`==1,1,0),
    `Protein` = ifelse(`Novavax NVX-CoV2373`==1 | `CIGB CIGB-66/Abdala`==1 | `Instituto Finlay de Vacunas Soberana 02`==1 | `AZLB ZF2001`==1 | `Vector Institute EpiVacCorona`==1 | 
                             `Medigen MVC-COV1901`==1 | `Instituto Finlay de Vacunas Soberana 01`==1 | `Biological E Corbevax`==1 | `Sanofi/GSK CoV2 preS dTM`==1 | `Medicago CoVLP`==1,1,0),
  )

# Save output
write.csv(Ineq_vaccine_coverage_platform, "Data/Processed/Ineq_merged_with_vaccine_data.csv")


#Pivot longer + plat classification
owid_vaccine_country_g <- Ineq_vaccine_coverage_platform %>% 
    pivot_longer(
    cols = `CIGB CIGB-66/Abdala`: `Valneva VLA2001`,
    names_to = "vaccine",
    values_to = "check",
    values_drop_na = TRUE) %>% 
  filter(
    check== 1
  ) %>% 
  select(-check) %>%
  mutate(
    vac_type = case_when(
      vaccine %in% c("BioNTech/Pfizer BNT162b2", "Moderna mRNA-1273") ~ "mRNA",
      vaccine %in% c("Oxford/AstraZeneca ChAdOx1-S", "Janssen Ad26.COV2.S", "Gamaleya Gam-COVID-Vac/Sputnik V", "Cansino Ad5-nCoV") ~ "Vectored",
      vaccine %in% c("Beijing/Sinopharm BBIBP-CorV", "Sinovac CoronaVac", "Bharat Covaxin/BBV152", "Wuhan/Sinopharm vaccine", "RIBSP Kazakhstan QazCovid-in", "Shifa Pharmed COVIran", "Erciyes TURKOVAC", "Shenzhen Kangtai KCONVAC", "Institute of Medical Biology CAMS vaccine", "Chumakov Center CoviVac", "Valneva VLA2001") ~ "Inactivated",
      vaccine %in% c("Novavax NVX-CoV2373", "CIGB CIGB-66/Abdala", "Instituto Finlay de Vacunas Soberana 02", "AZLB ZF2001", "Vector Institute EpiVacCorona", "Medigen MVC-COV1901", "Instituto Finlay de Vacunas Soberana 01", "Biological E Corbevax", "Sanofi/GSK CoV2 preS dTM", "Medicago CoVLP") ~ "Protein",
      TRUE ~ "Unknown"
    )
  ) 

unique(owid_vaccine_country_g$vaccine)

#Global vaccine platform %
global_vac_plat_distrib <-owid_vaccine_country_g %>% 
  select(
    Country.Code, q_GNI_per_c_WB, vac_type
  ) %>% 
  distinct() %>% 
  filter(
    !is.na(q_GNI_per_c_WB)
  ) %>%  
  group_by(vac_type) %>% 
  summarise(total=n()) %>% 
  filter(
    vac_type=="Inactivated" |vac_type=="mRNA" |vac_type=="Vectored" |vac_type=="Protein"
  ) %>% 
  mutate(
    vac_type_per = round(total / 194 * 100, 1),
    ) 

global_vac_plat_distrib




#Vaccine platform by GNI quartile
vac_plat_distrib <-owid_vaccine_country_g %>% 
  select(
    Country.Code, q_GNI_per_c_WB, vac_type
  ) %>% 
  distinct() %>% 
  filter(
    !is.na(q_GNI_per_c_WB)
  ) %>%  
  group_by(q_GNI_per_c_WB, vac_type) %>% 
  summarise(total=n()) %>% 
  filter(
    vac_type=="Inactivated" |vac_type=="mRNA" |vac_type=="Vectored" |vac_type=="Protein"
  ) %>% 
  mutate(
    countries_quartile=case_when(
      q_GNI_per_c_WB == 1 ~ 48,
      q_GNI_per_c_WB == 2 ~ 49,
      q_GNI_per_c_WB == 3 ~ 49,
      q_GNI_per_c_WB == 4 ~ 48
    )
  ) %>% 
  mutate(
    vac_type_per = round(total / countries_quartile * 100, 1),
    ) 

vac_plat_distrib


```

```{r}
vac_plat_distrib$vac_type <- factor(vac_plat_distrib$vac_type, levels = c("mRNA", "Vectored", "Inactivated", "Protein"))


d_vac_plat_distrib_plot<- ggplot(vac_plat_distrib, aes(x = as.factor(q_GNI_per_c_WB), y = vac_type_per, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()+
  facet_wrap(~ vac_type, ncol=4) +
  labs(title = "(D)",
       x = "",
       y = " % countries using platform (07/01/2024)") +
    guides(fill = "none") +  # Remove the fill legend
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  theme_minimal() +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )
d_vac_plat_distrib_plot

```


```{r}
vac_distrib <-owid_vaccine_country_g %>% 
  select(
    Country.Code, q_GNI_per_c_WB, vaccine
  ) %>% 
  group_by(vaccine) %>% 
  mutate(tot_vac = n()) %>% 
  distinct() %>% 
  filter(
    !is.na(q_GNI_per_c_WB)
  ) %>%  
  group_by(q_GNI_per_c_WB, vaccine, tot_vac) %>% 
  summarise(total_GNI=n())%>% 
  mutate(
    countries_quartile=case_when(
      q_GNI_per_c_WB == 1 ~ 48,
      q_GNI_per_c_WB == 2 ~ 49,
      q_GNI_per_c_WB == 3 ~ 49,
      q_GNI_per_c_WB == 4 ~ 48
    )
  ) %>% 
  mutate(
    vac_per = round(total_GNI / countries_quartile * 100, 1),
    ) 
```


```{r}
# Figure 1.
#Original
(map1_a / map1_b) |
  (c1 / c2_plot_intro_scale_delay / d_vac_plat_distrib_plot)


# Define areas
layout <- "
AAA
AAA
AAA
AAA
BBB
BBB
BBB
BBB
CCE
DDE
"

# Combine plots into layout
final_plot <- map1_a + 
              map1_b + 
              c1 + 
              d_vac_plat_distrib_plot + 
              c2_plot_intro_scale_delay +
              plot_layout(design = layout)

# Display final plot
final_plot

```




```{r Supplementary Figure 2}
# Supplementary Figure 2. Percentage of countries using different vaccine platforms  by GNI quartile

vac_name_distrib <-owid_vaccine_country_g %>% 
  select(
    Country.Code, q_GNI_per_c_WB, vac_type, vaccine
  ) %>% 
  distinct() %>% 
  filter(
    !is.na(q_GNI_per_c_WB)
  ) %>% 
  group_by(q_GNI_per_c_WB, vaccine, vac_type) %>% 
  summarise(total=n()) %>% 
  # filter(
  #   vac_type=="Inactivated" |vac_type=="mRNA" |vac_type=="Vectored"
  # ) %>%
  mutate(
    countries_quartile=case_when(
      q_GNI_per_c_WB == 1 ~ 48,
      q_GNI_per_c_WB == 2 ~ 49,
      q_GNI_per_c_WB == 3 ~ 49,
      q_GNI_per_c_WB == 4 ~ 48
    )
  ) %>% 
  mutate(
    vac_name_per = round(total / countries_quartile * 100, 1),
    ) 

vac_name_distrib

#Plot
vac_name_distrib$vac_type <- factor(vac_name_distrib$vac_type, levels = c("mRNA", "Vectored", "Inactivated", "Protein"))


vac_name_distrib <- vac_name_distrib %>%
  arrange(vac_type, -vac_name_per) %>%
  mutate(order_variable = row_number()) %>%
  ungroup()

# set vaccine factor order
vac_name_distrib$vaccine = factor(vac_name_distrib$vaccine, levels = unique(vac_name_distrib$vaccine))

s2_vac_name_distrib_plot<- ggplot(vac_name_distrib, aes(x = as.factor(q_GNI_per_c_WB), y = vac_name_per, fill = as.factor(vac_type))) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()+
  facet_wrap(~vaccine,ncol = 5 ) +
  labs(title = "",
       x = "GNI quartile",
       y = " % countries using vaccine",
       fill = "Vaccine Type") +
  scale_fill_manual(values = c("#EECC66", "#EE99AA", "#6699CC", "#009999"),
    labels = c("mRNA","Vectored", "Inactivated", "Protein")) +
  theme_minimal() +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )
  
s2_vac_name_distrib_plot
```



```{r}
#Possible supplementary figure

map_GNI_q <- ggplot(data = world_ineq_gral) +
  geom_sf(aes(fill = as.factor(q_GNI_per_c_WB)), position = "identity") + 
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866"),
                    labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)","no data\navailable")) +
  labs(
    fill = "GNI per capita quartile") +
  theme_void()+
  guides(color = guide_legend(title.position = "bottom", nrow = 1))+
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9),  # Adjust text size
    legend.title = element_text(size = 8),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )

map_GNI_q


```


##Fig 2 on the ‘Coverage over time in high-priority groups’?

<!-- (A) Coverage in HCWs by GNI as of (top) Dec 2021 and (bottom) Dec 2023 -->
<!-- (B) Coverage in older adults by GNI as of (top) Dec 2021 and (bottom) Dec 2023 -->
<!-- (C) Coverage per capita by GNI as of (top) Dec 2021 and (bottom) Dec 2023 -->

<!-- Each would be a box plot in the style of Fig 1C. Or something like that?  -->


```{r}
high_priority_metrics <- left_join(WHO_hcw_old_data, WHO_hcw_old_sum_data) %>% 
  arrange(date, iso) %>% 
  group_by(iso) %>% 
  mutate(
    vac_intro_hcw = ifelse(row_number() == which(cov_hcw_cps > 0)[1], as.character(date), NA),
    vac_F40_hcw = ifelse(row_number() == which(cov_hcw_cps > 0.3999)[1], as.character(date), NA),
    vac_F70_hcw = ifelse(row_number() == which(cov_hcw_cps > 0.6999)[1], as.character(date), NA),
    vac_intro_old = ifelse(row_number() == which(cov_old_cps > 0)[1], as.character(date), NA),
    vac_F40_old = ifelse(row_number() == which(cov_old_cps > 0.3999)[1], as.character(date), NA),
    vac_F70_old = ifelse(row_number() == which(cov_old_cps > 0.6999)[1], as.character(date), NA)
  ) %>%
  filter(!is.na(vac_F70_old) | !is.na(vac_F40_old) | !is.na(vac_intro_old) | !is.na(vac_F70_hcw) | !is.na(vac_F40_hcw) | !is.na(vac_intro_hcw)) %>% 
  select(!(iso:pol_age_old_source)) %>% 
  fill(vac_intro_hcw:vac_F70_old)%>%
  slice_tail() %>%  
  ungroup() %>% 
  mutate(
    V1_hcw_dummy = case_when(
      !is.na(vac_intro_hcw) & ymd(vac_intro_hcw) < ymd("2022-01-01") ~ "yes",
      is.na(vac_intro_hcw) ~ NA,
      TRUE ~ "no"
    ),
    V40_hcw_dummy = case_when(
      !is.na(vac_F40_hcw) & ymd(vac_F40_hcw) < ymd("2022-01-01") ~ "yes",
      is.na(vac_intro_hcw) ~ NA,
      TRUE ~ "no"
    ),
    V70_hcw_dummy = case_when(
      !is.na(vac_F70_hcw) & ymd(vac_F70_hcw) < ymd("2022-07-01") ~ "yes",
      is.na(vac_intro_hcw) ~ NA,
      TRUE ~ "no"
    ),
    V1_old_dummy = case_when(
      !is.na(vac_intro_old) & ymd(vac_intro_old) < ymd("2022-01-01") ~ "yes",
      is.na(vac_intro_old) ~ NA,
      TRUE ~ "no"
    ),
    V40_old_dummy = case_when(
      !is.na(vac_F40_old) & ymd(vac_F40_old) < ymd("2022-01-01") ~ "yes",
      is.na(vac_intro_old) ~ NA,
      TRUE ~ "no"
    ),
    V70_old_dummy = case_when(
      !is.na(vac_F70_old) & ymd(vac_F70_old) < ymd("2022-07-01") ~ "yes",
      is.na(vac_intro_old) ~ NA,
      TRUE ~ "no"
    ),
    V1_hcw_dummy_nd = case_when(
      !is.na(vac_intro_hcw) & ymd(vac_intro_hcw) < ymd("2024-01-01") ~ "yes",
      is.na(vac_intro_hcw) ~ NA,
      TRUE ~ "no"
    ),
    V40_hcw_dummy_nd = case_when(
      !is.na(vac_F40_hcw) & ymd(vac_F40_hcw) < ymd("2024-01-01") ~ "yes",
      is.na(vac_intro_hcw) ~ NA,
      TRUE ~ "no"
    ),
    V70_hcw_dummy_nd = case_when(
      !is.na(vac_F70_hcw) & ymd(vac_F70_hcw) < ymd("2024-01-01") ~ "yes",
      is.na(vac_intro_hcw) ~ NA,
      TRUE ~ "no"
    ),
    V1_old_dummy_nd = case_when(
      !is.na(vac_intro_old) & ymd(vac_intro_old) < ymd("2024-01-01") ~ "yes",
      is.na(vac_intro_old) ~ NA,
      TRUE ~ "no"
    ),
    V40_old_dummy_nd = case_when(
      !is.na(vac_F40_old) & ymd(vac_F40_old) < ymd("2024-01-01") ~ "yes",
      is.na(vac_intro_old) ~ NA,
      TRUE ~ "no"
    ),
    V70_old_dummy_nd = case_when(
      !is.na(vac_F70_old) & ymd(vac_F70_old) < ymd("2024-01-01") ~ "yes",
      is.na(vac_intro_old) ~ NA,
      TRUE ~ "no"
    )
  )

coverage_high_2021_2023<- WHO_hcw_old_data %>% 
  select(iso, date, cov_hcw_cps, cov_old_cps, cov_tot_cps, cov_tot_boost, cov_hcw_boost, cov_old_boost) %>% 
  filter(date=="2021-12-01" | date=="2023-12-01" | date=="2023-11-01") %>% 
  pivot_wider(., names_from = date, values_from = c(cov_hcw_cps, cov_old_cps, cov_tot_cps, cov_tot_boost, cov_hcw_boost, cov_old_boost))

Ineq_vaccine_coverage<- left_join(Ineq_vaccine_coverage, high_priority_metrics, by=c("iso3c"= "iso"))

Ineq_vaccine_coverage<- left_join(Ineq_vaccine_coverage, coverage_high_2021_2023, by=c("iso3c"= "iso"))

# Since some countries end reporting at 11/2023 rather than 12/2023, take maximum value 
Ineq_vaccine_coverage <- Ineq_vaccine_coverage %>%
  mutate(
    `cov_hcw_cps_2023-12-01` = if_else(!is.na(`cov_hcw_cps_2023-12-01`), `cov_hcw_cps_2023-12-01`, `cov_hcw_cps_2023-11-01`),
    `cov_old_cps_2023-12-01` = if_else(!is.na(`cov_old_cps_2023-12-01`), `cov_old_cps_2023-12-01`, `cov_old_cps_2023-11-01`),
    `cov_tot_cps_2023-12-01` = if_else(!is.na(`cov_tot_cps_2023-12-01`), `cov_tot_cps_2023-12-01`, `cov_tot_cps_2023-11-01`),
    `cov_hcw_boost_2023-12-01` = if_else(!is.na(`cov_hcw_boost_2023-12-01`), `cov_hcw_boost_2023-12-01`, `cov_hcw_boost_2023-11-01`),
    `cov_old_boost_2023-12-01` = if_else(!is.na(`cov_old_boost_2023-12-01`), `cov_old_boost_2023-12-01`, `cov_old_boost_2023-11-01`),
    `cov_tot_boost_2023-12-01` = if_else(!is.na(`cov_tot_boost_2023-12-01`), `cov_tot_boost_2023-12-01`, `cov_tot_boost_2023-11-01`)
  ) %>%
  select(-ends_with('2023-11-01'))
  
# Save output
write.csv(Ineq_vaccine_coverage, "Data/Processed/Ineq_merged_with_vaccine_data_plus_priority_groups.csv")
```

# Boxplots health care workers and older elderly


```{r}
#### Boxplot milestone December 2021/2023
collated = rbind(
  Ineq_vaccine_coverage %>% mutate(coverage = `cov_hcw_cps_2021-12-01`, date = "December 2021", group = "HCWs"),
  Ineq_vaccine_coverage %>% mutate(coverage = `cov_hcw_cps_2023-12-01`, date = "December 2023", group = "HCWs"),
  Ineq_vaccine_coverage %>% mutate(coverage = `cov_old_cps_2021-12-01`, date = "December 2021", group = "Older adults"),
  Ineq_vaccine_coverage %>% mutate(coverage = `cov_old_cps_2023-12-01`, date = "December 2023", group = "Older adults"),
  Ineq_vaccine_coverage %>% mutate(coverage = `cov_tot_cps_2021-12-01`, date = "December 2021", group = "Whole population"),
  Ineq_vaccine_coverage %>% mutate(coverage = `cov_tot_cps_2023-12-01`, date = "December 2023", group = "Whole population")
)

collated$group = factor(collated$group, levels=unique(collated$group))

ggplot(collated) +
  geom_boxplot(aes(x = coverage*100, y = as.factor(q_GNI_per_c_WB)), outlier.shape = NA) +
   geom_point(
    aes(
     x = coverage*100, 
     y = as.factor(q_GNI_per_c_WB),
      #size = studies,
      color = as.factor(q_GNI_per_c_WB)
    ),
    alpha = 0.5,
    position = position_jitterdodge(dodge.width = 0.5, jitter.width = 1) 
  )+
  scale_color_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  facet_grid(group~date)+
  labs(#title = "Health Care workers",
       x = "Primary series coverage (%)", 
       y = "GNI quartile")  +
  guides(color = "none") +
  scale_y_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal() + 
  theme(
    panel.spacing = unit(1.5, "lines"),
    legend.position = "bottom",
    legend.text = element_text(size = 8),# Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) # Adjust plot title size
  ) 
```


```{r}
# HCW

#median dates to built groups
summary_vacc_intro_scale_hcw <- Ineq_vaccine_coverage %>%
  mutate(
    vac_intro_hcw_month = floor_date(as.Date(vac_intro_hcw), "month"),
    vac_F40_hcw_month = floor_date(as.Date(vac_F40_hcw), "month"),
    vac_F70_hcw_month = floor_date(as.Date(vac_F70_hcw), "month")
  ) %>%
  group_by(q_GNI_per_c_WB) %>%
  summarize(
    n = length(q_GNI_per_c_WB),
    n_intro = sum(!is.na(vac_intro_hcw_month)),
    median_value_1 = format(as.Date(median(as.numeric(vac_intro_hcw_month), na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q25_1 = format(as.Date(quantile(as.numeric(vac_intro_hcw_month), 0.25, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q75_1 = format(as.Date(quantile(as.numeric(vac_intro_hcw_month), 0.75, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    median_value_40 = format(as.Date(median(as.numeric(vac_F40_hcw_month), na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q25_40 = format(as.Date(quantile(as.numeric(vac_F40_hcw_month), 0.25, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q75_40 = format(as.Date(quantile(as.numeric(vac_F40_hcw_month), 0.75, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    median_value_70 = format(as.Date(median(as.numeric(vac_F70_hcw_month), na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q25_70 = format(as.Date(quantile(as.numeric(vac_F70_hcw_month), 0.25, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q75_70 = format(as.Date(quantile(as.numeric(vac_F70_hcw_month), 0.75, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m")
  ) %>% 
  mutate(
    introduction = paste0( median_value_1, " (", q25_1, "-", q75_1, ")"),
    scale_40 = paste0( median_value_40, " (", q25_40, "-", q75_40, ")"),
    scale_70 = paste0( median_value_70, " (", q25_70, "-", q75_70, ")")
  ) %>% 
  select(q_GNI_per_c_WB, introduction, scale_40, scale_70) 


# Milestones
hcw_milestone_achieved <- Ineq_vaccine_coverage %>%
  group_by(q_GNI_per_c_WB) %>%
  summarise(
    V1 = sum(V1_hcw_dummy == "yes", na.rm = TRUE),
    V40 = sum(V40_hcw_dummy == "yes", na.rm = TRUE),
    V70 = sum(V70_hcw_dummy == "yes", na.rm = TRUE),
    V1_nd = sum(V1_hcw_dummy_nd == "yes", na.rm = TRUE),
    V40_nd = sum(V40_hcw_dummy_nd == "yes", na.rm = TRUE),
    V70_nd = sum(V70_hcw_dummy_nd == "yes", na.rm = TRUE),
    total = sum(V1_hcw_dummy_nd == "yes", na.rm = TRUE),
    V1_per = round(sum(V1_hcw_dummy == "yes", na.rm = TRUE) / total * 100, 0),
    V40_per = round(sum(V40_hcw_dummy == "yes", na.rm = TRUE) / total * 100, 0),
    V70_per = round(sum(V70_hcw_dummy == "yes", na.rm = TRUE) / total * 100, 0),
    V1_per_nd = round(sum(V1_hcw_dummy_nd == "yes", na.rm = TRUE) / total * 100, 0),
    V40_per_nd = round(sum(V40_hcw_dummy_nd == "yes", na.rm = TRUE) / total * 100, 0),
    V70_per_nd = round(sum(V70_hcw_dummy_nd == "yes", na.rm = TRUE) / total * 100, 0)
  ) %>% 
  mutate(
    V1_per_number = paste0(V1, "/", total, " (", V1_per, "%)"),
    V40_per_number = paste0(V40, "/", total, " (", V40_per, "%)"),
    V70_per_number = paste0(V70, "/", total, " (", V70_per, "%)"),
    V1_per_nd_number = paste0(V1_nd, "/", total, " (", V1_per_nd, "%)"),
    V40_per_nd_number = paste0(V40_nd, "/", total, " (", V40_per_nd, "%)"),
    V70_per_nd_number = paste0(V70_nd, "/", total, " (", V70_per_nd, "%)")
  ) %>% 
  select(q_GNI_per_c_WB, V1_per_number:V70_per_nd_number)


```




```{r}
# old

#median dates to built groups
summary_vacc_intro_scale_old <- Ineq_vaccine_coverage %>%
  mutate(
    vac_intro_old_month = floor_date(as.Date(vac_intro_old), "month"),
    vac_F40_old_month = floor_date(as.Date(vac_F40_old), "month"),
    vac_F70_old_month = floor_date(as.Date(vac_F70_old), "month")
  ) %>%
  group_by(q_GNI_per_c_WB) %>%
  summarize(
    n = length(q_GNI_per_c_WB),
    n_intro = sum(!is.na(vac_intro_old_month)),
    median_value_1 = format(as.Date(median(as.numeric(vac_intro_old_month), na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q25_1 = format(as.Date(quantile(as.numeric(vac_intro_old_month), 0.25, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q75_1 = format(as.Date(quantile(as.numeric(vac_intro_old_month), 0.75, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    median_value_40 = format(as.Date(median(as.numeric(vac_F40_old_month), na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q25_40 = format(as.Date(quantile(as.numeric(vac_F40_old_month), 0.25, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q75_40 = format(as.Date(quantile(as.numeric(vac_F40_old_month), 0.75, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    median_value_70 = format(as.Date(median(as.numeric(vac_F70_old_month), na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q25_70 = format(as.Date(quantile(as.numeric(vac_F70_old_month), 0.25, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m"),
    q75_70 = format(as.Date(quantile(as.numeric(vac_F70_old_month), 0.75, na.rm = TRUE), origin = "1970-01-01"), "%Y/%m")
  ) %>% 
  mutate(
    introduction = paste0( median_value_1, " (", q25_1, "-", q75_1, ")"),
    scale_40 = paste0( median_value_40, " (", q25_40, "-", q75_40, ")"),
    scale_70 = paste0( median_value_70, " (", q25_70, "-", q75_70, ")")
  ) %>% 
  select(q_GNI_per_c_WB, introduction, scale_40, scale_70) 


# Milestones
old_milestone_achieved <- Ineq_vaccine_coverage %>%
  group_by(q_GNI_per_c_WB) %>%
  summarise(
    V1 = sum(V1_old_dummy == "yes", na.rm = TRUE),
    V40 = sum(V40_old_dummy == "yes", na.rm = TRUE),
    V70 = sum(V70_old_dummy == "yes", na.rm = TRUE),
    V1_nd = sum(V1_old_dummy_nd == "yes", na.rm = TRUE),
    V40_nd = sum(V40_old_dummy_nd == "yes", na.rm = TRUE),
    V70_nd = sum(V70_old_dummy_nd == "yes", na.rm = TRUE),
    total = sum(V1_old_dummy_nd == "yes", na.rm = TRUE),
    V1_per = round(sum(V1_old_dummy == "yes", na.rm = TRUE) / total * 100, 0),
    V40_per = round(sum(V40_old_dummy == "yes", na.rm = TRUE) / total * 100, 0),
    V70_per = round(sum(V70_old_dummy == "yes", na.rm = TRUE) / total * 100, 0),
    V1_per_nd = round(sum(V1_old_dummy_nd == "yes", na.rm = TRUE) / total * 100, 0),
    V40_per_nd = round(sum(V40_old_dummy_nd == "yes", na.rm = TRUE) / total * 100, 0),
    V70_per_nd = round(sum(V70_old_dummy_nd == "yes", na.rm = TRUE) / total * 100, 0)
  ) %>% 
  mutate(
    V1_per_number = paste0(V1, "/", total, " (", V1_per, "%)"),
    V40_per_number = paste0(V40, "/", total, " (", V40_per, "%)"),
    V70_per_number = paste0(V70, "/", total, " (", V70_per, "%)"),
    V1_per_nd_number = paste0(V1_nd, "/", total, " (", V1_per_nd, "%)"),
    V40_per_nd_number = paste0(V40_nd, "/", total, " (", V40_per_nd, "%)"),
    V70_per_nd_number = paste0(V70_nd, "/", total, " (", V70_per_nd, "%)")
  ) %>% 
  select(q_GNI_per_c_WB, V1_per_number:V70_per_nd_number)


```


```{r}
#Overall coverage dec 2021/ 2023
summary_overall_cov_21_23 <- collated %>%
  select(iso3c, q_GNI_per_c_WB, `cov_hcw_cps_2023-12-01`:`cov_tot_cps_2021-12-01`) %>% 
  unique() %>% 
  filter(!is.na(iso3c)) %>% 
  group_by(q_GNI_per_c_WB) %>%
  summarize(
    n = length(q_GNI_per_c_WB),
    n_hcw_dec21 = sum(!is.na(`cov_hcw_cps_2021-12-01`)),
    n_hcw_dec23 = sum(!is.na(`cov_hcw_cps_2023-12-01`)),
    n_old_dec21 = sum(!is.na(`cov_old_cps_2021-12-01`)),
    n_old_dec23 = sum(!is.na(`cov_old_cps_2023-12-01`)),
    n_tot_dec21 = sum(!is.na(`cov_tot_cps_2021-12-01`)),
    n_tot_dec23 = sum(!is.na(`cov_tot_cps_2023-12-01`)),
    median_hcw_dec21 = round((median(as.numeric(`cov_hcw_cps_2021-12-01`), na.rm = TRUE)*100),0),
    q25_hcw_dec21 = round((quantile(as.numeric(`cov_hcw_cps_2021-12-01`), 0.25, na.rm = TRUE)*100),0),
    q75_hcw_dec21 = round((quantile(as.numeric(`cov_hcw_cps_2021-12-01`), 0.75, na.rm = TRUE)*100),0),
    median_hcw_dec23 = round((median(as.numeric(`cov_hcw_cps_2023-12-01`), na.rm = TRUE)*100),0),
    q25_hcw_dec23 = round((quantile(as.numeric(`cov_hcw_cps_2023-12-01`), 0.25, na.rm = TRUE)*100),0),
    q75_hcw_dec23 = round((quantile(as.numeric(`cov_hcw_cps_2023-12-01`), 0.75, na.rm = TRUE)*100),0),
    median_old_dec21 = round((median(as.numeric(`cov_old_cps_2021-12-01`), na.rm = TRUE)*100),0),
    q25_old_dec21 = round((quantile(as.numeric(`cov_old_cps_2021-12-01`), 0.25, na.rm = TRUE)*100),0),
    q75_old_dec21 = round((quantile(as.numeric(`cov_old_cps_2021-12-01`), 0.75, na.rm = TRUE)*100),0),
    median_old_dec23 = round((median(as.numeric(`cov_old_cps_2023-12-01`), na.rm = TRUE)*100),0),
    q25_old_dec23 = round((quantile(as.numeric(`cov_old_cps_2023-12-01`), 0.25, na.rm = TRUE)*100),0),
    q75_old_dec23 = round((quantile(as.numeric(`cov_old_cps_2023-12-01`), 0.75, na.rm = TRUE)*100),0),
    median_tot_dec21 = round((median(as.numeric(`cov_tot_cps_2021-12-01`), na.rm = TRUE)*100),0),
    q25_tot_dec21 = round((quantile(as.numeric(`cov_tot_cps_2021-12-01`), 0.25, na.rm = TRUE)*100),0),
    q75_tot_dec21 = round((quantile(as.numeric(`cov_tot_cps_2021-12-01`), 0.75, na.rm = TRUE)*100),0),
    median_tot_dec23 = round((median(as.numeric(`cov_tot_cps_2023-12-01`), na.rm = TRUE)*100),0),
    q25_tot_dec23 = round((quantile(as.numeric(`cov_tot_cps_2023-12-01`), 0.25, na.rm = TRUE)*100),0),
    q75_tot_dec23 = round((quantile(as.numeric(`cov_tot_cps_2023-12-01`), 0.75, na.rm = TRUE)*100),0)
  ) %>% 
  mutate(
    hcw_dec21 = paste0( median_hcw_dec21, "% (", q25_hcw_dec21, "%-", q75_hcw_dec21, "%)"),
    hcw_dec23 = paste0( median_hcw_dec23, "% (", q25_hcw_dec23, "%-", q75_hcw_dec23, "%)"),
    old_dec21 = paste0( median_old_dec21, "% (", q25_old_dec21, "%-", q75_old_dec21, "%)"),
    old_dec23 = paste0( median_old_dec23, "% (", q25_old_dec23, "%-", q75_old_dec23, "%)"),
    tot_dec21 = paste0( median_tot_dec21, "% (", q25_tot_dec21, "%-", q75_tot_dec21, "%)"),
    tot_dec23 = paste0( median_tot_dec23, "% (", q25_tot_dec23, "%-", q75_tot_dec23, "%)")
  ) %>% 
  select(q_GNI_per_c_WB:n_tot_dec23, hcw_dec21:tot_dec23) 
```


Supplementary scatter plot

```{r}

#Scatter plot + smooth line
# x= GNI + y = date
milestone_country <- Ineq_vaccine_coverage %>%
  select(Country.Name, GNI_per_c_WB, q_GNI_per_c_WB, min_vac_intro, owid_vac_F40, owid_vac_F70) %>%
  mutate(
    owid_vac_F40 = as.Date(owid_vac_F40), 
    owid_vac_F70 = as.Date(owid_vac_F70)
  ) %>% 
  pivot_longer(
    cols = c(min_vac_intro, owid_vac_F40, owid_vac_F70),
    names_to = "milestone",
    values_to = "date"
  )

  
plot_milestone_country <- ggplot(milestone_country) +
  geom_point(aes(x = log(GNI_per_c_WB), y = date, color = as.factor(q_GNI_per_c_WB))) +
  geom_smooth(aes(
    x = log(GNI_per_c_WB), y = date),
    color = "black", 
    linetype = "dotted", 
    size = 0.5,     
    se = TRUE,       
    alpha = 0.3)+       
  facet_wrap(~ milestone, 
             scales = "free", 
             labeller = labeller(milestone = c("min_vac_intro" = "Vaccine introduction", "owid_vac_F40" = "Vaccine scale-up (40%)","owid_vac_F70" = "Vaccine scale-up (70%)" ))) + 
  scale_y_date(
    limits = c(as.Date("2020-12-01"), NA),
               date_labels = "%b/%Y") +
  labs(
    title = "Vaccine Milestones by GNI per Capita",
    x = "GNI per capita (log)",
    y = "Date",
    color = "GNI quartile"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

plot_milestone_country

```

