---
title: "3 Creation of VE Database"
author: "Martina Pesce & Edward P. K. Parker"
date: "2023-10-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

#### Inputs
- Input/WeeklySummary_COVID19_VE_Studies_20230622.xlsx: View-Hub VE database
- Processed/Ineq_merged.csv: Combined country-level income data
- Input/meta_effect/VE_vaccine_list.csv: clean vaccine names and platforms for VE data
- Input/meta_effect/VE_variant_list.csv: clean variant names for VE data

#### Outputs
- Processed/COVID_VEcleaned.csv: Cleaned View-Hub VE data
- Processed/COVID19VE_ineq_merged.csv: View-Hub VE data merged with country income data


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
#libraries

library(readxl)
library(xlsx)
library(countrycode)
library(tidyverse)
library(lubridate)
```


```{r,import database}
#Vaccine Effectiveness Weekly studies database (11-01-2024)

COVID19VE <- read.xlsx("Data/Input/meta_effect/WeeklySummary_COVID19_VE_Studies_20240111_0.xlsx",sheetName = "Primary Series Studies", check.names=FALSE) 

# Update non-standard strings to avoid later issues
COVID19VE$timing_of_dose_days = str_replace_all(COVID19VE$timing_of_dose_days, "≤", ">=")
COVID19VE$timing_of_dose_days = str_replace_all(COVID19VE$timing_of_dose_days, "≥", ">=")
COVID19VE$population_detail = str_replace_all(COVID19VE$population_detail, "≥", ">=")
COVID19VE$population_category = str_replace_all(COVID19VE$population_category, "≥", ">=")
COVID19VE$`first author` = str_replace_all(COVID19VE$`first author`, "Šmíd", "Smid")
COVID19VE$`first author` = str_replace_all(COVID19VE$`first author`, "Á", "A")
COVID19VE$`first author` = str_replace_all(COVID19VE$`first author`, "á", "a")
COVID19VE$`first author` = str_replace_all(COVID19VE$`first author`, "ó", "o")

# Basic summary statistics
nrow(COVID19VE) #3732 estimates
length(unique(COVID19VE$`study ID`)) #434 unique studies

# Income data
Ineq_merged <- read_csv("Data/Processed/Ineq_merged_with_vaccine_data.csv") %>%
  select(`Country.Name`:`Protein`)

# Clean vaccine names, platforms, and variants for VE data
clean_vaccines <- read_csv("Data/Input/meta_effect/VE_vaccine_list.csv") 
clean_variants <- read_csv("Data/Input/meta_effect/VE_variant_list.csv") 
```

# View-Hub database

```{r}
#Explore country levels in VIEW-HUB database
table(COVID19VE$country)
```

Studies involving multiple countries include either commas or numerals, with the exception of 'Hong Kong, SAR'.

## Recode countries names

```{r harmonize country names}
COVID19VE <- COVID19VE %>%
  mutate(country = ifelse(grepl("Israel ", country), "Israel", country)) %>% # Merge multiple spellings of Israel
  mutate(country = ifelse(grepl("Hong Kong, SAR", country), "Hong Kong", country)) %>% # Update Hong Kong to remove comma 
  mutate(country_clean = ifelse(grepl("[0-9,]", as.character(country)) | grepl("countries", as.character(country)), "Multiple", country)) # recode list of countries to multiple country

# Additional cleaning steps
COVID19VE$country_clean[grepl("Canada", as.character(COVID19VE$country))] = "Canada"
COVID19VE$country_clean[grepl("Scotland", as.character(COVID19VE$country))] = "UK"
COVID19VE$country_clean[grepl("UAE", as.character(COVID19VE$country))] = "United Arab Emirates"
table(COVID19VE$country_clean)

# Add ISO3 codes for merge
COVID19VE <- COVID19VE %>%
  mutate(iso3c = countrycode(country_clean, origin = "country.name", destination = "iso3c"))

#Check iso3c NAs
filter <- COVID19VE %>% filter(is.na(iso3c))
all(filter$country_clean=="Multiple") # TRUE, so all filtered rows from multiple studies
length(unique(filter$`study ID`)) #11 unique studies removed

# Filter studies from multiple countries
COVID19VE_country <- COVID19VE %>% 
  filter(!is.na(iso3c)) %>%
  # Divide country-specific estimates into separate IDs
  mutate(
    #Study ID = (1)240 two countries -> create a new ID for 240+Scotland=500 / 222+Scotland=501/ (2) fix error in Tartof study ID
    Study_id_f = case_when(
      `study ID` == 240 & country == "Scotland" ~ 500,
      `study ID` == 222 & country == "Scotland" ~ 501,
      `first author` == "Tartof" & `publication date` == "22Apr2022" ~ 165,
      TRUE ~ `study ID`
    )
  )

# Basic summary statistics
nrow(COVID19VE_country) #3640 estimates
length(unique(COVID19VE_country$Study_id_f)) #425 unique studies

# Filter studies without income data
COVID19VE_country[!(COVID19VE_country$iso3c %in% Ineq_merged$iso3c),] # 3 studies - 1 from Cuba (Más-Bermejo), 1 from Taiwan (Huang) and 1 from Puerto Rico (Irizarry)
COVID19VE_country <- COVID19VE_country %>% filter(iso3c %in% Ineq_merged$iso3c)
nrow(COVID19VE_country) #3599 estimates
length(unique(COVID19VE_country$Study_id_f)) #423 unique studies
```


# Merge Systematic review with inequalities data.

```{r merge}
#Merge inequalities data with the VE data
# exclude product/platform binary variables from merge to avoid confusion with VE estimates
COVID19VE_ineq_merged <- left_join(COVID19VE_country, 
                                   Ineq_merged %>% select(Country.Code:V70_dummy), 
                                   by = "iso3c", na_matches = "never")
```



#Fix data problems
```{r}
# Write csv files to generate VE_vaccine_list.csv and VE_variant_list.csv inputs
#write.csv(unique(COVID19VE_ineq_merged$`primary series vaccine`),"vaccine_list.csv")
#write.csv(unique(COVID19VE_ineq_merged$variant_context),"variant_list.csv")
unique(COVID19VE_ineq_merged$outcome_category)

#fix a study that have 2 countries
COVID19VE_ineq_merged <- COVID19VE_ineq_merged %>%
  mutate(
    #Fix outcome 
    outcome_category = case_when(
    `study ID` == 345 &
        outcome_detail == "severe disease" ~ "severe",
      TRUE ~ outcome_category  
    ),
    #Fix publication dates
    `publication date` = case_when(
      `publication date` == "Jun2022 (day not specified)" ~ "01Jun2022",
      `publication date` == "Apr2022 (day not specified)" ~ "01Apr2022",
      Study_id_f == 102  ~ "05Jan2022",
      Study_id_f == 203 &
        `publication date` == "06Jul2022" ~ "05Nov2022",
      Study_id_f == 204  ~ "23May2022",
      Study_id_f == 235  ~ "22Jul2022",
      Study_id_f == 139  ~ "21Jan2022",
      TRUE ~ `publication date`
    ),
    #  publication date to dmy
    publication_date = dmy(paste0(
      substring(`publication date`, 1, 2),
      tolower(substring(`publication date`, 3, 5)),
      substring(`publication date`, 6)
    )),
  # Regroup outcome
  outcome_final = case_when(
    outcome_category %in% c(
    "asymptomatic",
    "infection",
    "infection ",
    "infection in the first 2 months of life",
    "infection in the first 4 months of life",
    "infection in the first 6 months of life",
    "infection during an infant's first 4 months of life",
    "reinfection"
  ) ~ "infection",
      outcome_category %in% c(
    "symptomatic",
    "symptomatic or severe" # grouped with less severe (more common) outcome
  ) ~ "symptomatic",
  outcome_category %in% c(
    "severe",
    "hospitalization",
    "hospitalization with MIS-C",
    "MIS-C"
  ) ~ "severe",
  outcome_category %in% c("death", "death ") ~ "death",
  TRUE ~ "X"
    )
  ) %>%
  #Create publication semester
  mutate(
    semester = case_when(
      publication_date < dmy("01-07-2021") ~ "2021 (Jan-Jun)",
      publication_date >= dmy("01-07-2021") & publication_date < dmy("01-01-2022") ~ "2021 (Jul-Dec)",
      publication_date >= dmy("01-01-2022") & publication_date < dmy("01-07-2022") ~ "2022 (Jan-Jun)",
      publication_date >= dmy("01-07-2022") & publication_date < dmy("01-01-2023") ~ "2022 (Jul-Dec)",
      publication_date >= dmy("01-01-2023") & publication_date < dmy("01-07-2023") ~ "2023 (Jan-Jun)",
      publication_date >= dmy("01-07-2023") & publication_date < dmy("01-01-2024") ~ "2023 (Jul-Dec)",
      publication_date >= dmy("01-01-2024") ~ "2024 (Jan-Jun)",
      TRUE ~ NA_character_
    )
  ) 

# add clean vaccine names
all(clean_vaccines$`primary series vaccine` %in% COVID19VE_ineq_merged$`primary series vaccine`)
COVID19VE_ineq_merged <- COVID19VE_ineq_merged %>% left_join(clean_vaccines, by=c("primary series vaccine")) 

# add clean variant names
all(clean_variants$variant_context %in% COVID19VE_ineq_merged$variant_context)
COVID19VE_ineq_merged <- COVID19VE_ineq_merged %>% left_join(clean_variants, by=c("variant_context"))

# Summarise
nrow(COVID19VE_ineq_merged) #3599 estimates
length(unique(COVID19VE_ineq_merged$Study_id_f)) #423 unique studies
```

```{r}
# earliest publication date
studies_min_pub_date <- read_csv("Data/Input/meta_effect/min_study_dates.csv") %>% 
  select(`study ID`, earliest_date) 

# Merge with COVID19VE_ineq_merged
COVID19VE_ineq_merged <- left_join(COVID19VE_ineq_merged, studies_min_pub_date) %>% 
  mutate(
    earliest_date= case_when(
    Study_id_f== "165" ~ "18/01/2022",
    TRUE ~ earliest_date
  ))



#Creat earliest publication date
COVID19VE_ineq_merged <- COVID19VE_ineq_merged %>% 
  mutate(
      min_pub_date = pmin(publication_date, dmy(earliest_date), na.rm = TRUE)
  )

# check

check1 <- COVID19VE_ineq_merged  %>% 
  filter(
    Study_id_f == "165"
  )%>% 
  select(
    Study_id_f, `first author`, `study ID`, publication_date, `publication date`, earliest_date, min_pub_date
  )

check <- COVID19VE_ineq_merged %>% 
  select(`study ID`, `first author`, country_clean, q_GNI_per_c_WB, `publication date`, publication_date, earliest_date, min_pub_date) %>% 
  distinct() %>% 
  mutate(
    dif_date= publication_date - min_pub_date
  )
sum(check$dif_date!=0)
#137 dates fixed 



```


```{r, write}
#Save data the .csv file for the descriptive analysis and metaregression

write.csv(COVID19VE_ineq_merged, "Data/Processed/COVID19VE_ineq_merged.csv")

```


  
# Session info
```{r, session info}
print(sessionInfo())
```

