---
title: "4 VE description (Fig 2)"
author: "Martina Pesce & Edward P. K. Parker"
date: "2023-11-30"
output: html_document
editor_options: 
  chunk_output_type: console
---


#### Inputs
- Processed/COVID19VE_ineq_merged.csv: View-Hub VE data merged with country income data
- Processed/Ineq_merged.csv: Combined country-level income data
- Input/Vaccine_data/owid-covid-data.csv: Our World in Data COVID-19 database
- Input/Vaccine_data/owid_vaccine_plat.csv: Our World in Data COVID-19 vaccine platforms
- Input/Vaccine_data/vaccination-data.csv: WHO COVID-19 vaccine database

#### Outputs
- Descriptive statistics on VE study distribution
- Figure 2


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r Libraries}
#Libraries
library(tidyverse)
library(readr)
library(rnaturalearth)
library(sf)
#library(viridisLite)
#library(viridis)
library(lubridate)
library(countrycode)
library(patchwork) 
library(cowplot)
```


```{r Data base}
#Databases

#Eff database
COVID19VE_ineq_merged <- read_csv("Data/Processed/COVID19VE_ineq_merged.csv") 

#Ineq database
Ineq_merged <- read_csv("Data/Processed/Ineq_merged_with_vaccine_data.csv") 

```

###########################################################################################################################################################
## Fig 1. maps: Global inequities in the evaluation of vaccine effectiveness. (A) Map of the distribution of available VE studies across different countries.## 
###########################################################################################################################################################




```{r data for map}
#Create table with study counts and introduction dates by country
Country_studies <- COVID19VE_ineq_merged %>%
  group_by(Country.Code, country_clean) %>%
  summarise(
    min_VE_pub_date = min(as.Date(min_pub_date)),
    n_VE_studies = n_distinct(Study_id_f),
    perc_VE_studies = round(n_distinct(Study_id_f)/423*100,0)
  )
nrow(Country_studies) # 48 distinct countries with country-specific VE estimates
nstudies = sum(Country_studies$n_VE_studies) 
nstudies # 422 unique studies
Country_studies %>% arrange(-n_VE_studies) %>% filter(n_VE_studies>10)

# Check all country iso codes represented in inequity data
all(Country_studies$Country.Code %in% Ineq_merged$Country.Code)

#Merge and create study categories
Ineq_merged <- left_join(Ineq_merged, Country_studies, by="Country.Code") %>% 
  mutate(
      study_dummy = case_when(
      !is.na(n_VE_studies) ~ "yes",
      TRUE ~ "no"
    )) %>%
  # add time to publication metrics
   mutate(
    diff_V40_VEpub = as.Date(min_VE_pub_date) - as.Date(owid_vac_F40),
    diff_intro_VE_pub = as.Date(min_VE_pub_date) - as.Date(min_vac_intro)
    )

#Geographical information

countries <-  ne_countries(returnclass = "sf") %>% 
  mutate(iso3c = countrycode(admin, origin = "country.name", destination = "iso3c")) %>% 
  filter(iso3c!="ATA")
  

world_ineq_gral <- 
  merge(countries, Ineq_merged, by.x = "iso3c", by.y = "Country.Code", all = TRUE) %>% 
    mutate(
    n_VE_studies_g = case_when(
      n_VE_studies > 0 & n_VE_studies < 6 ~ "01-05",
      n_VE_studies >= 6 & n_VE_studies < 11 ~ "06-10",
      n_VE_studies >= 11 & n_VE_studies < 16 ~ "11-15",
      n_VE_studies >= 16 ~ "16+",
      TRUE ~ "0"
    )
    )
  
world_ineq_gral <- st_as_sf(world_ineq_gral)

```


```{r}
#Map A: vaccine introduction dates

map1_a<- ggplot(data = world_ineq_gral) +
  geom_sf(aes(fill = as.factor(V1_date_group)), position = "identity") + 
  scale_fill_manual(values = c("#2e73b8","#75b2d8","#9dd5ea", "#d7e0f4","#A2B1A4"),
                    labels = c("<Jan\n2021", "Jan\n2021", "Feb\n2021", ">Mar\n2021", "not\nreached"),
                    breaks = c("before jan/2021", "jan/2021", "feb/2021", "march/2021 +", "not reached"),
                    na.value = "#F0F3F0") +
  labs(
    title= "(A)\nVaccine introduction date",
    fill = "Month") +
  theme_void()+
  guides(color = guide_legend(title.position = "bottom", nrow = 1))+
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9),  # Adjust text size
    legend.title = element_text(size = 8),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )

map1_a

#Map 1.B: scale-up dates
map1_b<- ggplot(data = world_ineq_gral) +
  geom_sf(aes(fill = as.factor(V40_date_group)), position = "identity") + 
  scale_fill_manual(values = c("#827717","#afb42b", "#cddc39","#ecebb1","#A2B1A4"),
                    breaks = c("before Jun/2021", "Jun-Aug/2021", "Sep-Nov/2021", "Dec/2021 +", "not reached"),
                    labels = c("<Jun\n2021", "Jun-Aug\n2021", "Sep-Nov\n2021", ">Dec\n2021", "not\nreached"),
                    na.value = "#F0F3F0") +
  labs(
           title= "(B)\nTime of vaccine scale-up (40%)",
    fill = "Month") +
  theme_void()+ 
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 9),  # Adjust text size
    legend.title = element_text(size = 8),  # Adjust legend title size
    plot.title = element_text(size = 10) # Adjust plot title size
  ) 

map1_b

#Map 1.C: number of VE studies by country

map1_c<- ggplot(data = world_ineq_gral) +
  geom_sf(aes(fill = n_VE_studies_g), 
          position = "identity") + 
  scale_fill_manual(values = c("#F0F0F0", "#C8E6A4", "#79C77A", "#3D8B62", "#005147"),
                    labels = c("0","01-05", "06-10", "11-15", "16+"))+ 
  labs(
    title= "(C)",
    fill = 'Number of\nVE studies') +
  theme_void() +
  theme(
    legend.position = "bottom",        # Move legend to the bottom
    legend.key.width = unit(0.5, "cm"), # Adjust the width of legend key to make it smaller
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )

map1_c


map_1 <- map1_a / map1_b / map1_c
map_1


```


#######################################
# Figure 2
#############

```{r}

## Fig 2.a:  Percentage of countries achieving vaccine introduction (left) and scale-up to 40% primary series coverage by 31 December 2021 (right) by GNIpc

summary_milestone_achieved <- Ineq_merged %>% 
  group_by(q_GNI_per_c_WB) %>% 
  summarise(
    V1_per = round(sum(V1_dummy == "yes", na.rm = TRUE) / n() * 100, 1),
    V40_per = round(sum(V40_dummy == "yes", na.rm = TRUE) / 
                    sum(V1_dummy == "yes", na.rm = TRUE) * 100, 1),
        study_per = round(sum(study_dummy == "yes")/ sum(V1_dummy == "yes", na.rm = TRUE) * 100, 1))

summary_milestone_achieved

summary_milestone_achieved_l <- summary_milestone_achieved %>% 
    pivot_longer(
    cols = c(V1_per, V40_per, study_per),
    names_to = "milestone",
    values_to = "Value"
  )

summary_milestone_achieved_l$milestone <- factor(
  summary_milestone_achieved_l$milestone,
  levels = c("V1_per", "V40_per", "study_per")
)
summary_milestone_achieved_l

# Plot
Fig_2_a_vac_achiev <- ggplot(summary_milestone_achieved_l, aes(x = as.factor(q_GNI_per_c_WB), y = Value, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  coord_flip()+
facet_wrap(~ milestone, labeller = labeller(milestone = c(
  "V1_per" = "Vaccine introduction (31/12/2021)", 
  "V40_per" = "Vaccine scale-up (40%, 31/12/2021)",
  "study_per" = "% countries with ≥1 VE studies (07/01/2024)"
))) +
  labs(title = "(A)",
       x = "",
       y = "% countries achieving milestone") +
    guides(fill = "none") +  # Remove the fill legend
  theme_minimal() +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal() + 
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),# Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) # Adjust plot title size
  ) 

Fig_2_a_vac_achiev



```


```{r}
#Fig 2.b:  Time to vaccine introduction (left) and scale-up (right) by GNIpc

#Data for plot
intro_scale_delay<- Ineq_merged %>%
  mutate(
    owid_vac_F40 = as.Date(owid_vac_F40)  ) %>% 
  select(Country.Name, q_GNI_per_c_WB, min_vac_intro, owid_vac_F40, min_VE_pub_date) %>%
  filter(!is.na(min_vac_intro)) %>%
  pivot_longer(
    cols = c(min_vac_intro, owid_vac_F40, min_VE_pub_date),
    names_to = "Date",
    values_to = "Value"
  ) 

# Set factor levels
 intro_scale_delay$Date = factor(intro_scale_delay$Date, levels = unique(intro_scale_delay$Date))

#Plot
Fig_2_b_vac_delay <- ggplot(intro_scale_delay) +
  geom_boxplot(aes(x = Value, y = as.factor(q_GNI_per_c_WB)), outlier.shape = NA) +
   geom_point(
    aes(
     x = Value, 
     y = as.factor(q_GNI_per_c_WB),
      #size = studies,
      color = as.factor(q_GNI_per_c_WB)
    ),
    alpha = 0.5,
    position = position_jitterdodge(dodge.width = 0.5, jitter.width = 1) 
  )+
  scale_x_date(limits = c(as.Date("2020-12-01"), NA),
               date_labels = "%b/%Y") +
  scale_color_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  facet_wrap(~Date, scales = "free_x", labeller = labeller(Date = c("min_vac_intro" = "Date of vaccine introduction", "owid_vac_F40" = "Date of vaccine scale-up (40%)", "min_VE_pub_date"= "Date of first VE study")), nrow=1)+
  labs(title = "(B)",
       x = "Date", 
       y = "GNI quartile")  +
  guides(color = "none") +
  scale_y_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal() + 
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),# Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) # Adjust plot title size
  ) 
 
Fig_2_b_vac_delay 
```



```{r}
#Pivot longer + plat classification
owid_vaccine_country_g <- Ineq_merged %>% 
    pivot_longer(
    cols = `CIGB CIGB-66/Abdala`: `Valneva VLA2001`,
    names_to = "vaccine",
    values_to = "check",
    values_drop_na = TRUE) %>% 
  filter(
    check== 1
  ) %>% 
  select(-check) %>%
  mutate(
    vac_type = case_when(
      vaccine %in% c("BioNTech/Pfizer BNT162b2", "Moderna mRNA-1273") ~ "mRNA",
      vaccine %in% c("Oxford/AstraZeneca ChAdOx1-S", "Janssen Ad26.COV2.S", "Gamaleya Gam-COVID-Vac/Sputnik V", "Cansino Ad5-nCoV") ~ "Vectored",
      vaccine %in% c("Beijing/Sinopharm BBIBP-CorV", "Sinovac CoronaVac", "Bharat Covaxin/BBV152", "Wuhan/Sinopharm vaccine", "RIBSP Kazakhstan QazCovid-in", "Shifa Pharmed COVIran", "Erciyes TURKOVAC", "Shenzhen Kangtai KCONVAC", "Institute of Medical Biology CAMS vaccine", "Chumakov Center CoviVac", "Valneva VLA2001") ~ "Inactivated",
      vaccine %in% c("Novavax NVX-CoV2373", "CIGB CIGB-66/Abdala", "Instituto Finlay de Vacunas Soberana 02", "AZLB ZF2001", "Vector Institute EpiVacCorona", "Medigen MVC-COV1901", "Instituto Finlay de Vacunas Soberana 01", "Biological E Corbevax", "Sanofi/GSK CoV2 preS dTM", "Medicago CoVLP") ~ "Protein",
      TRUE ~ "Unknown"
    )
  ) 

unique(owid_vaccine_country_g$vaccine)

#Vaccine platform by GNI quartile
summary_vac_plat_distrib_GNI <-owid_vaccine_country_g %>% 
  select(
    Country.Code, q_GNI_per_c_WB, vac_type
  ) %>% 
  distinct() %>% 
  filter(
    !is.na(q_GNI_per_c_WB)
  ) %>%  
  group_by(q_GNI_per_c_WB, vac_type) %>% 
  summarise(total=n()) %>% 
  filter(
    vac_type=="Inactivated" |vac_type=="mRNA" |vac_type=="Vectored" |vac_type=="Protein"
  ) %>% 
  mutate(
    countries_quartile=case_when(
      q_GNI_per_c_WB == 1 ~ 50,
      q_GNI_per_c_WB == 2 ~ 51,
      q_GNI_per_c_WB == 3 ~ 51,
      q_GNI_per_c_WB == 4 ~ 51
    )
  ) %>% 
  mutate(
    vac_type_per = round(total / countries_quartile * 100, 1),
    ) 

summary_vac_plat_distrib_GNI

summary_vac_plat_distrib_GNI$vac_type <- factor(summary_vac_plat_distrib_GNI$vac_type, levels = c("mRNA", "Vectored", "Inactivated", "Protein"))

# Plot
Fig_2_c_plat <- ggplot(summary_vac_plat_distrib_GNI, aes(x = as.factor(q_GNI_per_c_WB), y = vac_type_per, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()+
  facet_wrap(~ vac_type, ncol=4) +
  labs(title = "(C)",
       x = "",
       y = " % countries using platform (07/01/2024)") +
    guides(fill = "none") +  # Remove the fill legend
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  theme_minimal() +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )

Fig_2_c_plat
```



#??? here????
```{r}

#Create table with study counts and introduction dates by GNI quartile
GNI_studies <- COVID19VE_ineq_merged %>%
  group_by(q_GNI_per_c_WB) %>%
  summarise(
    min_VE_pub_date = min(as.Date(min_pub_date)),
    n_VE_studies = n_distinct(Study_id_f),
    perc_VE_studies = round(n_distinct(Study_id_f)/423*100,0)
  )

#Create table with study counts and introduction dates by vaccine platform and outcome
plat_outcome_studies <- COVID19VE_ineq_merged %>%
  group_by(vac_type, outcome_final) %>%
  summarise(
    min_VE_pub_date = min(as.Date(min_pub_date)),
    n_VE_studies = n_distinct(Study_id_f),
    perc_VE_studies = round(n_distinct(Study_id_f)/423*100,0)
  )
```




```{r}
# Percentage of VE studies by GNIpc GNI quartile.  
summ_GNI_eff_study <- Ineq_merged %>%
  select(Country.Name,q_GNI_per_c_WB, n_VE_studies, min_VE_pub_date) %>% 
  filter(!is.na(n_VE_studies)) %>%  # Exclude rows with NA in the 'n_VE_studies' variable
  group_by(
    q_GNI_per_c_WB
  ) %>% 
  summarize(
    total_st_GNI = sum(n_VE_studies),
    perc_st_GNI= sum(n_VE_studies)/422*100, #fix with final value
    min_VE_pub_date= min(min_VE_pub_date)
  )
summ_GNI_eff_study

# Plot (B)
plot2_b <- ggplot(summ_GNI_eff_study, aes(x = as.factor(q_GNI_per_c_WB), y = perc_st_GNI, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  coord_flip()+
  labs(title = "(B)",
       x = "GNI quartile",
       y = "% VE studies") +
    guides(fill = "none") +  # Remove the fill legend
  theme_minimal() +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)"))+
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10)
  ) +
  scale_y_continuous(limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))
plot2_b
```

```{r}
# (C) Percentage of countries with VE studies available by GNIpc GNI quartile.
Ineq_merged_introduced <- subset(Ineq_merged,  V1_dummy=="yes")

table(Ineq_merged_introduced$q_GNI_per_c_WB)

eff_study_perc <-Ineq_merged %>%
  group_by(q_GNI_per_c_WB) %>%
  summarise(
    study = sum(study_dummy == "yes"))%>%
  mutate(
    countries_quartile=case_when(
      q_GNI_per_c_WB == 1 ~ 50,
      q_GNI_per_c_WB == 2 ~ 51,
      q_GNI_per_c_WB == 3 ~ 51,
      q_GNI_per_c_WB == 4 ~ 51
    ),
  ) %>% 
  mutate(
    study_per = round(study / countries_quartile * 100, 0),
    )
eff_study_perc

#Plot (C)
plot2_c<- ggplot(eff_study_perc, aes(x = as.factor(q_GNI_per_c_WB), y = study_per, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  coord_flip()+ labs(title = "(C)",
       x = "",
       y = "% countries with VE studies") +
    guides(fill = "none") +  # Remove the fill legend
  theme_minimal() +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)"))+
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10)
  )  +
  scale_y_continuous(limits = c(0, 60))

plot2_c
```


```{r}
# (D) Time from vaccine introduction to first VE study publication in the country by GNIpc GNI quartile. Note that only countries with at least one available VE study are included
 plot2_d<-ggplot(Ineq_merged) +
  geom_boxplot(aes(x = diff_intro_VE_pub, y = as.factor(q_GNI_per_c_WB)), outlier.shape=NA) +
   geom_point(
    aes(
     x = diff_intro_VE_pub, y = as.factor(q_GNI_per_c_WB),
      #size = studies,
      color = as.factor(q_GNI_per_c_WB)
    ),
    alpha = 0.5,
    position = position_jitterdodge(dodge.width = 0.5, jitter.width = 1) 
  )+
  scale_color_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866"))+
  labs(
    title = "(D)",
    x = "Number of days from introduction to first VE publication", 
    y = "GNI quartile")  +
  guides(color = FALSE) +
  scale_y_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal()  +
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) 
   ) +
   scale_x_continuous(limits = c(0, 1050), breaks = c(0, 250, 500, 750, 1000)) 

plot2_d

summary_vacc_eff_delay <- Ineq_merged %>%
  group_by(q_GNI_per_c_WB) %>%
  summarize(
    median_value_1 = round(median(as.numeric(diff_intro_VE_pub), na.rm = TRUE),0),
    q25_1 = round(quantile(as.numeric(diff_intro_VE_pub), 0.25, na.rm = TRUE),0),
    q75_1 = round(quantile(as.numeric(diff_intro_VE_pub), 0.75, na.rm = TRUE),0)
  )
summary_vacc_eff_delay
```



```{r}
# (E) Percentage of VE studies by platform. The ‘Multiple’ category includes one comparative VE study involving the subunit vaccine NVX-CoV2373. Protein subunit with 1 effectiveness vaccine study was excluded

plat_summary <- COVID19VE_ineq_merged %>% 
  select(Study_id_f, min_pub_date, vac_type, q_GNI_per_c_WB) %>% 
  distinct() %>% 
  group_by(vac_type) %>% 
  summarise(
    min_pub_vac_type = min(as.Date(min_pub_date)),
    total_papers_vac_type= n(),
    per_papers_vac_type = n()/423*100
  )

plat_summary$vac_type <- factor(plat_summary$vac_type, levels = c("Protein subunit","Multiple", "Inactivated","Vectored","mRNA"))

#Plot
Fig_2_d_stu_plat<- ggplot(plat_summary, aes(x = vac_type, y =per_papers_vac_type , fill = vac_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()+
  labs(title = "(D)",
       x = "",
       y = "% VE studies") +
    guides(fill = "none") +  # Remove the fill legend
  theme_minimal() +
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10)
  ) +
  scale_fill_manual(values = c("#0072B2","#009E73", "#D55E00","#CC79A7", "#F0E442"))+
  scale_x_discrete(labels = c("Protein","Multiple", "Inactivated", "Vectored", "mRNA")) +
  scale_y_continuous(limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))

Fig_2_d_stu_plat

```

```{r}
#(F) Percentage of countries with VE studies relating to different vaccine platforms by GNIpc GNI quartile. VE, vaccine effectiveness. Protein subunit with 1 effectiveness vaccine study was excluded
vac_plat_distrib <- COVID19VE_ineq_merged %>% 
  select(Country.Code, q_GNI_per_c_WB, vac_type) %>% 
  distinct() %>%  
  group_by(q_GNI_per_c_WB, vac_type) %>% 
  summarise(total=n()) %>% 
  mutate(
    countries_quartile=case_when(
      q_GNI_per_c_WB == 1 ~ 50,
      q_GNI_per_c_WB == 2 ~ 51,
      q_GNI_per_c_WB == 3 ~ 51,
      q_GNI_per_c_WB == 4 ~ 51
    )
  ) %>% 
  mutate(
    vac_type_per = round(total / countries_quartile * 100, 0),
    ) 
vac_plat_distrib

#Plot

vac_plat_distrib$vac_type <- factor(vac_plat_distrib$vac_type, levels = c("Protein subunit", "mRNA", "Vectored", "Inactivated", "Multiple"))


plot2_f<- ggplot(vac_plat_distrib[vac_plat_distrib$vac_type != "Protein subunit", ], aes(x = as.factor(q_GNI_per_c_WB), y = vac_type_per, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()+
  facet_wrap(~ vac_type, nrow = 2) +
  labs(title = "(F)",
       x = "GNI quartile",
       y = "% countries with VE studies") +
    guides(fill = "none") +  # Remove the fill legend
scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal()+
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )+
  scale_y_continuous(limits = c(0, 50), breaks = c(0, 25, 50))
plot2_f

```


```{r}
# (fig2_map + plot2_d +
#   plot_layout(ncol = 1, heights = c(3, 1))) |
#   # Arrange plot2_b, plot2_c, plot2_e, and plot2_f together
#   ((plot2_b | plot2_c) /
#    (plot2_e | plot2_f))
# 
# ((fig2_map / plot2_d) |
#   (plot2_b / plot2_e) |
#    (plot2_c / plot2_f) +
#   plot_layout(nrow = 1, widths = c(3, 1, 1))) 
# 
# combined_plot <- (
#   (fig2_map / plot2_d) |
#   (plot2_b / plot2_e) |
#   (plot2_c / plot2_f)
# )
# 
# combined_plot + plot_layout(ncol = 3, widths = c(3, 1, 1), heights = c(3,1) )

# 
# col1<-(fig2_map + plot2_d +
#   plot_layout(ncol = 1, heights = c(1, 1)))
# 
# col2<-(plot2_b | plot2_c) /
#    (plot2_e | plot2_f)
# 
# combined_plot <- col1 | col2
# 
# combined_plot + plot_layout(ncol = 3, widths = c(3, 3, 3))
# 
# fig2_map
# plot2_d
# (plot2_b | plot2_c | plot2_e | plot2_f)
# Fig3_plot <- 
#   fig2_map /
#   (plot2_d | plot2_b | plot2_c) /
#   (plot2_e | plot2_f) +
#   plot_layout(heights = c(3, 1, 1))
# Fig3_plot

```


```{r}
# FIg 2
Fig_2_l3 <- (Fig_2_c_plat | Fig_2_d_stu_plat) + plot_layout(widths = c(2, 1))
Fig_2 <- (Fig_2_a_vac_achiev / Fig_2_b_vac_delay / Fig_2_l3)


```




```{r}
outcome_summary <- COVID19VE_ineq_merged %>% 
  select(Study_id_f, outcome_final) %>% 
  distinct() %>% 
  group_by(outcome_final) %>% 
  summarise(
    total_papers_out= n(),
    per_papers_out = round(n()/nstudies*100,0)
  )
outcome_summary
```

```{r}
variant_summary <- COVID19VE_ineq_merged %>% 
  select(Study_id_f, variant_clean) %>% 
  distinct() %>% 
  group_by(variant_clean) %>% 
  summarise(
    total_papers_out= n(),
    per_papers_out = round(n()/nstudies*100,0)
  )
variant_summary
```



#Supplementary figure

```{r F2}

# Figure 3.A Vaccine effectiveness papers temporal distribution by countries quartile economic inequalities distribution between countries (GNI per capita).  (Upper) Each point represents a paper according to the publication date by GNI per capita quartile. (Lower) Total published papers per year half according to the GNI per capita quartile. Vaccine effectiveness papers included in the 23 June 2023 View-Hub Primary Series Studies database.

data_F2 <- COVID19VE_ineq_merged %>% 
  select(country_clean, iso3c, min_pub_date, semester,Study_id_f, q_GNI_per_c_WB) %>% 
  distinct()

time_a<- data_F2 %>%
  ggplot(aes(x = as.Date(min_pub_date), y = as.factor(q_GNI_per_c_WB))) +
  geom_boxplot() +
  geom_jitter(aes(color = as.factor(q_GNI_per_c_WB), alpha = 0.5)) +
   guides(alpha = "none") +
  scale_color_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) + #I dont like the scale
  labs(x = "Publication Date", y = "GNI quartile")  +
  guides(color = FALSE) +
  scale_y_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal()

paper_sem_quart <- data_F2 %>%
  group_by(semester, q_GNI_per_c_WB) %>%
  summarise(
    total_papers = n()
  ) 
  
   
time_b <- ggplot(paper_sem_quart, aes(x = semester, y = total_papers, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity") +
  labs(x = "", y = "N VE studies", fill = "GNI quartile") +
  theme_minimal() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866"),
    labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)") ) 

```


## B


```{r }

# Figure 3.B Vaccine effectiveness papers temporal distribution by vaccine type.  (Upper) Each point represents a paper according to the publication date by vaccine type. (Lower) Total published papers per year half according to vaccine type. Vaccine effectiveness papers included in the 23 June 2023 View-Hub Primary Series Studies database.

data_F3 <- COVID19VE_ineq_merged %>% 
  select(country_clean, iso3c, min_pub_date, semester, Study_id_f, q_GNI_per_c_WB, vac_type) %>% 
  distinct() 

data_F3$vac_type <- factor(data_F3$vac_type, levels = c("Protein subunit","Multiple", "Inactivated", "Vectored", "mRNA"))


# (B) vaccine platform. In the upper panels, each point represents a study. 
timeB_a<- data_F3 %>%
  ggplot(aes(x = as.Date(min_pub_date), y = as.factor(vac_type))) +
  geom_boxplot() +
  geom_jitter(aes(color = as.factor(vac_type),alpha = 0.5)) +
  guides(alpha = "none") +
  scale_color_manual(values = c("#0072B2","#009E73", "#D55E00","#CC79A7", "#F0E442"))+
  labs(x = "Publication Date", y = "Vaccine platform")  +
  guides(color = FALSE) +
  scale_y_discrete(labels = c("Protein","Multiple", "Inactivated", "Vectored", "mRNA")) + 
  theme_minimal()

paper_sem_quart <- data_F3 %>%
  group_by(semester, vac_type) %>%
  summarise(
    total_papers = n()
  ) 
  
  
timeB_b <- ggplot(paper_sem_quart, aes(x = semester, y = total_papers, fill = as.factor(vac_type))) +
  geom_bar(stat = "identity") +
  labs(x = "", y = "N VE studies", fill = "Vaccine platform") +
  theme_minimal() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("#0072B2","#009E73", "#D55E00","#CC79A7", "#F0E442"),
    labels = c("Protein","Multiple", "Inactivated", "Vectored", "mRNA"))


# Extended Data Fig. 3. Distribution of vaccine effectiveness papers over time.
combined_plotB <- time_a/time_b/timeB_a/timeB_b
combined_plotB 



vac_type_sum2 <- data_F3 %>%
  group_by(vac_type) %>%
  summarise(min_date = min(as.Date(min_pub_date)),
            ) 



median_labels <- data_F3 %>%
  group_by(vac_type) %>%
  summarize(
    median_value = as.Date(median(as.numeric(min_pub_date), na.rm = TRUE), origin = "1970-01-01"),
    q25 = as.Date(quantile(as.numeric(min_pub_date), 0.25, na.rm = TRUE), origin = "1970-01-01"),
    q75 = as.Date(quantile(as.numeric(min_pub_date), 0.75, na.rm = TRUE), origin = "1970-01-01")
  )
```


#### Descriptive table (Supplement)


```{r}
#Supplementary table 2

# vaccine study distribution
table_vacc <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(clean_vaccine, total_study) %>%
  mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(clean_vaccine, `n(%)`) %>% 
  distinct()

table_vac2 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type, clean_vaccine) %>% 
  group_by(vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(vac_type, clean_vaccine,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, clean_vaccine, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = vac_type, values_from = `n(%)`)
  
combined_Vac <- left_join(table_vacc, table_vac2)


# vaccine platform study distribution
table_plat_vac <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(vac_type, total_study) %>%
  mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, `n(%)`) %>% 
  distinct()


#variant of interest
table_VoI1 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, variant_clean) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(variant_clean, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(variant_clean, `n(%)`) %>% 
  distinct()

table_VoI2 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type, variant_clean) %>% 
  group_by(vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(vac_type, variant_clean,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, variant_clean, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = vac_type, values_from = `n(%)`)
  
combined_VoI <- left_join(table_VoI1, table_VoI2)

#outcome
table_outcome1 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, outcome_final) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(outcome_final, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(outcome_final, `n(%)`) %>% 
  distinct()
table_outcome2 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type, outcome_final) %>% 
  group_by(vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(vac_type, outcome_final,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, outcome_final, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = vac_type, values_from = `n(%)`)
  
combined_outcome <- left_join(table_outcome1, table_outcome2)

#GNI
table_GNI1 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, q_GNI_per_c_WB) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(q_GNI_per_c_WB, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(q_GNI_per_c_WB, `n(%)`) %>% 
  distinct()
table_GNI2 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type, q_GNI_per_c_WB) %>% 
  group_by(vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(vac_type, q_GNI_per_c_WB,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, q_GNI_per_c_WB, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = vac_type, values_from = `n(%)`)
  
combined_GNI <- left_join(table_GNI1, table_GNI2)
```

#### Clean supplementary dataset
```{r}
supp_data <- Ineq_merged %>%
  select(
    Country.Name:source_website,
    mRNA:Protein,
    min_VE_pub_date:diff_intro_VE_pub
  ) %>%
  select(
    -c(Economy, `Income group`, perc_VE_studies, study_dummy, 
       Region, Country.Code, owid_who, diff_V40_VEpub),
    -starts_with("V1"), -starts_with("V4"), -starts_with("V7")
  ) %>%
  relocate(iso3c, .after = Country.Name) %>%
  relocate(continent, .after = iso3c) %>%
  relocate(q_GNI_per_c_WB, .after = GNI_per_c_WB) %>%
  rename(country = Country.Name,
         iso_code = iso3c,
         worldbank_GNI_per_capita = GNI_per_c_WB,
         worldbank_GNI_quartile = q_GNI_per_c_WB,
         worldbank_GNI_year = year_GNI,
         owid_vaccines = vaccines,
         owid_last_observation_date = last_observation_date,
         any_mRNA = mRNA,
         any_vectored = Vectored,
         any_inactivated = Inactivated,
         any_protein = Protein,
         owid_source_name = source_name,
         owid_source_website = source_website,
         ivac_min_VE_pub_date = min_VE_pub_date,
         ivac_n_VE_studies = n_VE_studies,
         diff_intro_min_VE_pub = diff_intro_VE_pub) %>%
  relocate(diff_intro_min_VE_pub, .after = ivac_n_VE_studies)




write.csv(supp_data, "Data/Processed/supplementary_data.csv")
```



```{r}

#Scatter plot + smooth line
# x= GNI + y = date
scatter_milestone_country <- Ineq_merged %>%
  select(Country.Name, GNI_per_c_WB, q_GNI_per_c_WB, min_vac_intro, owid_vac_F40, min_VE_pub_date) %>%
  mutate(
    owid_vac_F40 = as.Date(owid_vac_F40)
  ) %>% 
  pivot_longer(
    cols = c(min_vac_intro, owid_vac_F40, min_VE_pub_date),
    names_to = "milestone",
    values_to = "date"
  )

  
ED_F0_scatterplot_GNI_milestone <- ggplot(scatter_milestone_country) +
  geom_point(aes(x = log(GNI_per_c_WB), y = date, color = as.factor(q_GNI_per_c_WB))) +
  geom_smooth(aes(
    x = log(GNI_per_c_WB), y = date),
    color = "black", 
    linetype = "dotted", 
    size = 0.5,     
    se = TRUE,       
    alpha = 0.3)+       
  facet_wrap(~ milestone, 
             scales = "free", 
             labeller = labeller(milestone = c("min_vac_intro" = "Vaccine introduction", "owid_vac_F40" = "Vaccine scale-up (40%)" , "min_VE_pub_date"= "min_VE_pub_date"))) + 
  scale_y_date(
    limits = c(as.Date("2020-12-01"), NA),
               date_labels = "%b/%Y") +
  labs(
    title = "Vaccine Milestones by GNI per Capita",
    x = "GNI per capita (log)",
    y = "Date",
    color = "GNI quartile"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )+
  scale_color_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866"),labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)"))

ED_F0_scatterplot_GNI_milestone

#Correlation test: for adding to the legend
spearman_corr <- scatter_milestone_country %>%
  group_by(milestone) %>%
  summarise(
    cor = cor.test(log(GNI_per_c_WB), as.numeric(date), method = "spearman")$estimate
  )

spearman_corr
```

