---
title: "4 VE description (Fig 2)"
author: "Martina Pesce & Edward P. K. Parker"
date: "2023-11-30"
output: html_document
editor_options: 
  chunk_output_type: console
---


#### Inputs
- Processed/COVID19VE_ineq_merged.csv: View-Hub VE data merged with country income data
- Processed/Ineq_merged.csv: Combined country-level income data
- Input/Vaccine_data/owid-covid-data.csv: Our World in Data COVID-19 database
- Input/Vaccine_data/owid_vaccine_plat.csv: Our World in Data COVID-19 vaccine platforms
- Input/Vaccine_data/vaccination-data.csv: WHO COVID-19 vaccine database

#### Outputs
- Descriptive statistics on VE study distribution
- Figure 2


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r Libraries}
#Libraries
library(tidyverse)
library(readr)
library(rnaturalearth)
library(sf)
library(viridisLite)
library(viridis)
library(lubridate)
library(countrycode)
library(patchwork) 
library(cowplot)
```


```{r Data base}
#Databases

#Eff database
COVID19VE_ineq_merged <- read_csv("Data/Processed/COVID19VE_ineq_merged.csv") 

#Ineq database
Ineq_merged <- read_csv("Data/Processed/Ineq_merged_with_vaccine_data.csv") 
```

# Panel 1. 

## A. 

### Data wrangling

```{r data wrangling}
#Create table with study counts and introduction dates by country
Country_studies <- COVID19VE_ineq_merged %>%
  group_by(Country.Code, country_clean) %>%
  summarise(
    min_VE_pub_date = min(as.Date(min_pub_date)),
    n_VE_studies = n_distinct(Study_id_f),
    perc_VE_studies = round(n_distinct(Study_id_f)/422*100,0)
  )
nrow(Country_studies) # 48 distinct countries with country-specific VE estimates
nstudies = sum(Country_studies$n_VE_studies) 
nstudies # 422 unique studies
Country_studies %>% arrange(-n_VE_studies) %>% filter(n_VE_studies>10)

#Create table with study counts and introduction dates by GNI quartile
GNI_studies <- COVID19VE_ineq_merged %>%
  group_by(q_GNI_per_c_WB) %>%
  summarise(
    min_VE_pub_date = min(as.Date(min_pub_date)),
    n_VE_studies = n_distinct(Study_id_f),
    perc_VE_studies = round(n_distinct(Study_id_f)/422*100,0)
  )


# Check all country iso codes represented in inequity data
all(Country_studies$Country.Code %in% Ineq_merged$Country.Code)

#Merge and create study categories
Ineq_merged <- left_join(Ineq_merged, Country_studies, by="Country.Code") %>% 
  mutate(
      study_dummy = case_when(
      !is.na(n_VE_studies) ~ "yes",
      TRUE ~ "no"
    )) %>%
  # add time to publication metrics
   mutate(
    diff_V40_VEpub = as.Date(min_VE_pub_date) - as.Date(owid_vac_F40),
    diff_intro_VE_pub = as.Date(min_VE_pub_date) - as.Date(min_vac_intro)
    )

#Geographical information

countries <-  ne_countries(returnclass = "sf") %>% 
  mutate(iso3c = countrycode(admin, origin = "country.name", destination = "iso3c")) %>% 
  filter(iso3c!="ATA")
  

world_ineq_gral <- 
  merge(countries, Ineq_merged, by.x = "iso3c", by.y = "Country.Code", all = TRUE) %>% 
    mutate(
    n_VE_studies_g = case_when(
      n_VE_studies > 0 & n_VE_studies < 6 ~ "01-05",
      n_VE_studies >= 6 & n_VE_studies < 11 ~ "06-10",
      n_VE_studies >= 11 & n_VE_studies < 16 ~ "11-15",
      n_VE_studies >= 16 ~ "16+",
      TRUE ~ "0"
    )
    )
  

world_ineq_gral <- st_as_sf(world_ineq_gral)
```


```{r}
# paper distribution(iv)
fig2_map<- ggplot(data = world_ineq_gral) +
  geom_sf(aes(fill = n_VE_studies_g), 
          position = "identity") + 
  scale_fill_manual(values = c("#DDDDDD","#d9e8bd","#8cba9e","#338c73","#006666" ),
                    labels = c("0","01-05", "06-10", "11-15", "16+"))+ 
  labs(
    title= "(A)",
    fill = 'Number of\nVE studies') +
  theme_void() +
  theme(
    legend.position = "bottom",        # Move legend to the bottom
    legend.key.width = unit(0.5, "cm"), # Adjust the width of legend key to make it smaller
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )

fig2_map
```

```{r}
#Graph 2.b: % publications by GNI
summ_GNI_eff_study <- Ineq_merged %>%
  select(Country.Name,q_GNI_per_c_WB, n_VE_studies, min_VE_pub_date) %>% 
  filter(!is.na(n_VE_studies)) %>%  # Exclude rows with NA in the 'n_VE_studies' variable
  group_by(
    q_GNI_per_c_WB
  ) %>% 
  summarize(
    total_st_GNI = sum(n_VE_studies),
    perc_st_GNI= sum(n_VE_studies)/422*100, #fix with final value
    min_VE_pub_date= min(min_VE_pub_date)
  )
summ_GNI_eff_study

# Plot (B)
plot2_b <- ggplot(summ_GNI_eff_study, aes(x = as.factor(q_GNI_per_c_WB), y = perc_st_GNI, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  coord_flip()+
  labs(title = "(B)",
       x = "GNI quartile",
       y = "% VE studies") +
    guides(fill = "none") +  # Remove the fill legend
  theme_minimal() +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)"))+
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10)
  ) +
  scale_y_continuous(limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))
plot2_b
```

```{r}
#Graph 2.c: bar-plot countries with publication
Ineq_merged_introduced <- subset(Ineq_merged,  V1_dummy=="yes")

table(Ineq_merged_introduced$q_GNI_per_c_WB)

eff_study_perc <-Ineq_merged %>%
  group_by(q_GNI_per_c_WB) %>%
  summarise(
    study = sum(study_dummy == "yes"))%>%
  mutate(
    countries_quartile=case_when(
      q_GNI_per_c_WB == 1 ~ 48,
      q_GNI_per_c_WB == 2 ~ 49,
      q_GNI_per_c_WB == 3 ~ 49,
      q_GNI_per_c_WB == 4 ~ 48
    ),
  ) %>% 
  mutate(
    study_per = round(study / countries_quartile * 100, 0),
    )
eff_study_perc

#Plot (C)
plot2_c<- ggplot(eff_study_perc, aes(x = as.factor(q_GNI_per_c_WB), y = study_per, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  coord_flip()+ labs(title = "(C)",
       x = "",
       y = "% countries with VE studies") +
    guides(fill = "none") +  # Remove the fill legend
  theme_minimal() +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)"))+
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10)
  )  +
  scale_y_continuous(limits = c(0, 60))

plot2_c
```


```{r}
# Plot time to publication (D)
 plot2_d<-ggplot(Ineq_merged) +
  geom_boxplot(aes(x = diff_intro_VE_pub, y = as.factor(q_GNI_per_c_WB)), outlier.shape=NA) +
   geom_point(
    aes(
     x = diff_intro_VE_pub, y = as.factor(q_GNI_per_c_WB),
      #size = studies,
      color = as.factor(q_GNI_per_c_WB)
    ),
    alpha = 0.5,
    position = position_jitterdodge(dodge.width = 0.5, jitter.width = 1) 
  )+
  scale_color_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866"))+
  labs(
    title = "(D)",
    x = "Number of days from introduction to first VE publication", 
    y = "GNI quartile")  +
  guides(color = FALSE) +
  scale_y_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal()  +
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) 
   ) +
   scale_x_continuous(limits = c(0, 1050), breaks = c(0, 250, 500, 750, 1000)) 

plot2_d

summary_vacc_eff_delay <- Ineq_merged %>%
  group_by(q_GNI_per_c_WB) %>%
  summarize(
    median_value_1 = round(median(as.numeric(diff_intro_VE_pub), na.rm = TRUE),0),
    q25_1 = round(quantile(as.numeric(diff_intro_VE_pub), 0.25, na.rm = TRUE),0),
    q75_1 = round(quantile(as.numeric(diff_intro_VE_pub), 0.75, na.rm = TRUE),0)
  )
summary_vacc_eff_delay
```

```{r}
n_distinct(COVID19VE_ineq_merged$Study_id_f)
plat_summary <- COVID19VE_ineq_merged %>% 
  select(Study_id_f, min_pub_date, vac_type, q_GNI_per_c_WB) %>% 
  distinct() %>% 
  group_by(vac_type) %>% 
  summarise(
    min_pub_vac_type = min(as.Date(min_pub_date)),
    total_papers_vac_type= n(),
    per_papers_vac_type = n()/422*100
  )
plat_summary$vac_type <- factor(plat_summary$vac_type, levels = c("Multiple", "Inactivated","Vectored","mRNA"))
#Plot
plot2_e<- ggplot(plat_summary, aes(x = vac_type, y =per_papers_vac_type , fill = vac_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()+
  labs(title = "(E)",
       x = "",
       y = "% VE studies") +
    guides(fill = "none") +  # Remove the fill legend
  theme_minimal() +
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10)
  ) +
  scale_fill_manual(values = c("grey", "#6699CC","#EE99AA", "#EECC66"))+
  scale_x_discrete(labels = c("Multiple", "Inactivated", "Vectored", "mRNA")) +
  scale_y_continuous(limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))
plot2_e

```

```{r}

vac_plat_distrib <- COVID19VE_ineq_merged %>% 
  select(Country.Code, q_GNI_per_c_WB, vac_type) %>% 
  distinct() %>%  
  group_by(q_GNI_per_c_WB, vac_type) %>% 
  summarise(total=n()) %>% 
  mutate(
    countries_quartile=case_when(
      q_GNI_per_c_WB == 1 ~ 48,
      q_GNI_per_c_WB == 2 ~ 49,
      q_GNI_per_c_WB == 3 ~ 49,
      q_GNI_per_c_WB == 4 ~ 48
    )
  ) %>% 
  mutate(
    vac_type_per = round(total / countries_quartile * 100, 0),
    ) 
vac_plat_distrib

#Plot

vac_plat_distrib$vac_type <- factor(vac_plat_distrib$vac_type, levels = c("mRNA", "Vectored", "Inactivated", "Multiple"))


plot2_f<- ggplot(vac_plat_distrib, aes(x = as.factor(q_GNI_per_c_WB), y = vac_type_per, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()+
  facet_wrap(~ vac_type, nrow = 2) +
  labs(title = "(F)",
       x = "GNI quartile",
       y = "% countries with VE studies") +
    guides(fill = "none") +  # Remove the fill legend
scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) +
  scale_x_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal()+
  theme(
    legend.text = element_text(size = 8),  # Adjust text size
    legend.title = element_text(size = 9),  # Adjust legend title size
    plot.title = element_text(size = 10) 
  )+
  scale_y_continuous(limits = c(0, 50), breaks = c(0, 25, 50))
plot2_f

```

```{r}
outcome_summary <- COVID19VE_ineq_merged %>% 
  select(Study_id_f, outcome_final) %>% 
  distinct() %>% 
  group_by(outcome_final) %>% 
  summarise(
    total_papers_out= n(),
    per_papers_out = round(n()/nstudies*100,0)
  )
outcome_summary
```

```{r}
variant_summary <- COVID19VE_ineq_merged %>% 
  select(Study_id_f, variant_clean) %>% 
  distinct() %>% 
  group_by(variant_clean) %>% 
  summarise(
    total_papers_out= n(),
    per_papers_out = round(n()/nstudies*100,0)
  )
variant_summary
```


```{r}
# (fig2_map + plot2_d +
#   plot_layout(ncol = 1, heights = c(3, 1))) |
#   # Arrange plot2_b, plot2_c, plot2_e, and plot2_f together
#   ((plot2_b | plot2_c) /
#    (plot2_e | plot2_f))
# 
# ((fig2_map / plot2_d) |
#   (plot2_b / plot2_e) |
#    (plot2_c / plot2_f) +
#   plot_layout(nrow = 1, widths = c(3, 1, 1))) 
# 
# combined_plot <- (
#   (fig2_map / plot2_d) |
#   (plot2_b / plot2_e) |
#   (plot2_c / plot2_f)
# )
# 
# combined_plot + plot_layout(ncol = 3, widths = c(3, 1, 1), heights = c(3,1) )


col1<-(fig2_map + plot2_d +
  plot_layout(ncol = 1, heights = c(1, 1)))

col2<-(plot2_b | plot2_c) /
   (plot2_e | plot2_f)

combined_plot <- col1 | col2

combined_plot + plot_layout(ncol = 3, widths = c(3, 3, 3))

```


#Supplementary figure

```{r F2}

# Figure 3.A Vaccine effectiveness papers temporal distribution by countries quartile economic inequalities distribution between countries (GNI per capita).  (Upper) Each point represents a paper according to the publication date by GNI per capita quartile. (Lower) Total published papers per year half according to the GNI per capita quartile. Vaccine effectiveness papers included in the 23 June 2023 View-Hub Primary Series Studies database.

data_F2 <- COVID19VE_ineq_merged %>% 
  select(country_clean, iso3c, min_pub_date, semester,Study_id_f, q_GNI_per_c_WB) %>% 
  distinct()

time_a<- data_F2 %>%
  ggplot(aes(x = as.Date(min_pub_date), y = as.factor(q_GNI_per_c_WB))) +
  geom_boxplot() +
  geom_jitter(aes(color = as.factor(q_GNI_per_c_WB), alpha = 0.5)) +
   guides(alpha = "none") +
  scale_color_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866")) + #I dont like the scale
  labs(x = "Publication Date", y = "GNI quartile")  +
  guides(color = FALSE) +
  scale_y_discrete(labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)")) + 
  theme_minimal()

paper_sem_quart <- data_F2 %>%
  group_by(semester, q_GNI_per_c_WB) %>%
  summarise(
    total_papers = n()
  ) 
  
   
time_b <- ggplot(paper_sem_quart, aes(x = semester, y = total_papers, fill = as.factor(q_GNI_per_c_WB))) +
  geom_bar(stat = "identity") +
  labs(x = "", y = "N VE studies", fill = "GNI quartile") +
  theme_minimal() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("#77AADD","#44BB99","#AAAA00","#EE8866"),
    labels = c("1 (lower\nincome)", "2", "3", "4 (higher\nincome)") ) 

```


## B


```{r }

# Figure 3.B Vaccine effectiveness papers temporal distribution by vaccine type.  (Upper) Each point represents a paper according to the publication date by vaccine type. (Lower) Total published papers per year half according to vaccine type. Vaccine effectiveness papers included in the 23 June 2023 View-Hub Primary Series Studies database.

data_F3 <- COVID19VE_ineq_merged %>% 
  select(country_clean, iso3c, min_pub_date, semester, Study_id_f, q_GNI_per_c_WB, vac_type) %>% 
  distinct() 

data_F3$vac_type <- factor(data_F3$vac_type, levels = c("Multiple", "Inactivated", "Vectored", "mRNA"))

timeB_a<- data_F3 %>%
  ggplot(aes(x = as.Date(min_pub_date), y = as.factor(vac_type))) +
  geom_boxplot() +
  geom_jitter(aes(color = as.factor(vac_type),alpha = 0.5)) +
  guides(alpha = "none") +
  scale_color_manual(values = c("grey", "#6699CC","#EE99AA", "#EECC66"))+
  labs(x = "Publication Date", y = "Vaccine platform")  +
  guides(color = FALSE) +
  scale_y_discrete(labels = c("Multiple", "Inactivated", "Vectored", "mRNA")) + 
  theme_minimal()

paper_sem_quart <- data_F3 %>%
  group_by(semester, vac_type) %>%
  summarise(
    total_papers = n()
  ) 
  
   
timeB_b <- ggplot(paper_sem_quart, aes(x = semester, y = total_papers, fill = as.factor(vac_type))) +
  geom_bar(stat = "identity") +
  labs(x = "", y = "N VE studies", fill = "Vaccine platform") +
  theme_minimal() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("grey", "#6699CC","#EE99AA", "#EECC66"),
    labels = c("Multiple", "Inactivated", "Vectored", "mRNA"))

combined_plotB <- time_a/time_b/timeB_a/timeB_b
combined_plotB 
vac_type_sum2 <- data_F3 %>%
  group_by(vac_type) %>%
  summarise(min_date = min(as.Date(min_pub_date)),
            ) 



median_labels <- data_F3 %>%
  group_by(vac_type) %>%
  summarize(
    median_value = as.Date(median(as.numeric(min_pub_date), na.rm = TRUE), origin = "1970-01-01"),
    q25 = as.Date(quantile(as.numeric(min_pub_date), 0.25, na.rm = TRUE), origin = "1970-01-01"),
    q75 = as.Date(quantile(as.numeric(min_pub_date), 0.75, na.rm = TRUE), origin = "1970-01-01")
  )
```


#### Descriptive table (Supplement)


```{r}
#Supplementary table 2

# vaccine study distribution
table_vacc <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(clean_vaccine, total_study) %>%
  mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(clean_vaccine, `n(%)`) %>% 
  distinct()

table_vac2 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type, clean_vaccine) %>% 
  group_by(vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(vac_type, clean_vaccine,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, clean_vaccine, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = vac_type, values_from = `n(%)`)
  
combined_Vac <- left_join(table_vacc, table_vac2)


# vaccine platform study distribution
table_plat_vac <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(vac_type, total_study) %>%
  mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, `n(%)`) %>% 
  distinct()


#variant of interest
table_VoI1 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, variant_clean) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(variant_clean, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(variant_clean, `n(%)`) %>% 
  distinct()

table_VoI2 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type, variant_clean) %>% 
  group_by(vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(vac_type, variant_clean,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, variant_clean, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = vac_type, values_from = `n(%)`)
  
combined_VoI <- left_join(table_VoI1, table_VoI2)

#outcome
table_outcome1 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, outcome_final) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(outcome_final, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(outcome_final, `n(%)`) %>% 
  distinct()
table_outcome2 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type, outcome_final) %>% 
  group_by(vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(vac_type, outcome_final,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, outcome_final, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = vac_type, values_from = `n(%)`)
  
combined_outcome <- left_join(table_outcome1, table_outcome2)

#GNI
table_GNI1 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, q_GNI_per_c_WB) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(q_GNI_per_c_WB, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(q_GNI_per_c_WB, `n(%)`) %>% 
  distinct()
table_GNI2 <- COVID19VE_ineq_merged %>%
  distinct(Study_id_f, vac_type, q_GNI_per_c_WB) %>% 
  group_by(vac_type) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(vac_type, q_GNI_per_c_WB,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0( n, " (", round(n / total_study * 100, 0),"%)" )) %>% 
  ungroup() %>% 
  select(vac_type, q_GNI_per_c_WB, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = vac_type, values_from = `n(%)`)
  
combined_GNI <- left_join(table_GNI1, table_GNI2)
```

#### Clean supplementary dataset
```{r}
supp_data <- Ineq_merged %>%
  select(
    Country.Name:source_website,
    mRNA:Protein,
    min_VE_pub_date:diff_intro_VE_pub
  ) %>%
  select(
    -c(Economy, `Income group`, perc_VE_studies, study_dummy, 
       Region, Country.Code, owid_who, diff_min_vac_intro_1Dec2020, diff_V40_VEpub),
    -starts_with("V1"), -starts_with("V4"), -starts_with("V7")
  ) %>%
  relocate(iso3c, .after = Country.Name) %>%
  relocate(continent, .after = iso3c) %>%
  relocate(q_GNI_per_c_WB, .after = GNI_per_c_WB) %>%
  rename(country = Country.Name,
         iso_code = iso3c,
         worldbank_GNI_per_capita = GNI_per_c_WB,
         worldbank_GNI_quartile = q_GNI_per_c_WB,
         worldbank_GNI_year = year_GNI,
         owid_vaccines = vaccines,
         owid_last_observation_date = last_observation_date,
         diff_min_vac_intro_1Dec2020 = diff_min_vac_intro_1Dec2020_clean,
         any_mRNA = mRNA,
         any_vectored = Vectored,
         any_inactivated = Inactivated,
         any_protein = Protein,
         owid_source_name = source_name,
         owid_source_website = source_website,
         ivac_min_VE_pub_date = min_VE_pub_date,
         ivac_n_VE_studies = n_VE_studies,
         diff_intro_min_VE_pub = diff_intro_VE_pub) %>%
  relocate(diff_min_vac_intro_1Dec2020, .after = diff_owid_who_intro)   %>%
  relocate(diff_intro_min_VE_pub, .after = ivac_n_VE_studies)
write.csv(supp_data, "Data/Processed/supplementary_data.csv")
```
