---
title: "5 Meta-regression inclusion criteria"
author: "Martina Pesce & Edward P. K. Parker"
date: "2023-10-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

#### Inputs
- Processed/COVID19VE_ineq_merged.csv: merged VE/income data
- Input/meta_effect/population_included.xlsx: database of eligible population groups
- Input/meta_effect/VE_population_list.csv: population groups eligible for inclusion

#### Outputs
- Final/VEmetareg_data.csv: Database of study groups eligible for meta-regression


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, data and lib}

# Libraries
library(tidyverse)
library(readr)
library(readxl)
library(xlsx)


# Data
VEmetareg_data <- read_csv("Data/Processed/COVID19VE_ineq_merged.csv") %>%
  select(`study ID`:variant_clean, -owid_who)
VE_populations <- read_csv("Data/Input/meta_effect/VE_population_list.csv") 

```

#### Selection of potentially VE estimates

```{r, inclusion criteria}
#Studies populations included in View-Hub primary series
nrow(VEmetareg_data)  #3599 estimates
n_distinct(VEmetareg_data$Study_id_f) #423 studies

## Step 1: exclude studies of infection
VEmetareg_data <- VEmetareg_data %>% 
  filter(outcome_final=="severe" | outcome_final=="death") 
nrow(VEmetareg_data)  #1643 estimates
n_distinct(VEmetareg_data$Study_id_f) #250 studies

## Step 2: exclude non-eligible populations (including studies restricted to previously infected)
# write csv files to generate VE_population_list.csv inputs
#write.csv(unique(VEmetareg_data$population_category), "populations.csv")
table(VEmetareg_data$historyofCOVID)


# check all population categories included, then merge
all(VE_populations$population_category %in% VEmetareg_data$population_category)
VEmetareg_data <- VEmetareg_data %>% left_join(VE_populations, by=c("population_category")) 

#Excluded:
excluded_pop <- VEmetareg_data %>% 
  filter(include_population!="yes" |
        historyofCOVID == "confirmed prior infection" |
        historyofCOVID == "Omicron prior infection" |
        historyofCOVID == "pre-Omicron prior infection" |
        historyofCOVID == "previously infected only"
        ) %>% 
  select(include_population, inclusion_notes, historyofCOVID, Study_id_f) %>% 
  unique()


# filter
VEmetareg_data <- VEmetareg_data %>% 
  filter(include_population=="yes" &
        historyofCOVID != "confirmed prior infection" &
        historyofCOVID != "Omicron prior infection" &
        historyofCOVID != "pre-Omicron prior infection" &
        historyofCOVID != "previously infected only"
        )
nrow(VEmetareg_data)  #1231 estimates
n_distinct(VEmetareg_data$Study_id_f) #178 studies

included_studies<-VEmetareg_data %>% 
  select(Study_id_f, include_population) %>% 
  unique() 

excluded_studies<- left_join(excluded_pop, included_studies, by= "Study_id_f") %>% 
  filter(is.na(include_population.y))
#table(excluded_studies$inclusion_notes)



## Step 3: complete, vaccine-specific primary series of any COVID-19 vaccine
table(VEmetareg_data$dose)
table(VEmetareg_data$clean_vaccine)

VEmetareg_data <- VEmetareg_data %>%
  filter( dose == "final" & 
    clean_vaccine != "Multiple or heterologous RNA" &
    clean_vaccine != "Heterologous platforms")
nrow(VEmetareg_data) #835
n_distinct(VEmetareg_data$Study_id_f) #130

## Step 4: cohort/case-control design
table(VEmetareg_data$`study design`)
VEmetareg_data <- VEmetareg_data %>%
  filter(
    `study design` != "Screening method"
    )
nrow(VEmetareg_data) #834
n_distinct(VEmetareg_data$Study_id_f) #129

## Step 5: variant-specific estimates
table(VEmetareg_data$variant_clean)

VEmetareg_data <- VEmetareg_data %>%
  filter(
    variant_clean != "delta/omicron" &
    variant_clean != "other combinations"
    )

nrow(VEmetareg_data) #486
n_distinct(VEmetareg_data$Study_id_f) #76

## Add vaccine - dose count combination
VEmetareg_data$clean_vaccine_dose = paste0(VEmetareg_data$clean_vaccine," x ",VEmetareg_data$dose_number) 

## Step 6: exclude products with <10 unique studies
table(VEmetareg_data$clean_vaccine_dose)
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Beijing/Sinopharm BBIBP-CorV x 2")$Study_id_f) # 6
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Bharat Covaxin/BBV152 x 2")$Study_id_f) # 3
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="BioNTech/Pfizer BNT162b2 x 2")$Study_id_f) # 53
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Cansino Ad5-nCoV x 1")$Study_id_f) # 2
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Gamaleya Gam-COVID-Vac/Sputnik V x 2")$Study_id_f) # 6
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Janssen Ad26.COV2.S x 2")$Study_id_f) # 2
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Janssen Ad26.COV2.S x 1")$Study_id_f) # 13
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Moderna mRNA-1273 x 2")$Study_id_f) # 27
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Oxford/AstraZeneca ChAdOx1-S x 2")$Study_id_f) # 28
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Sinovac CoronaVac x 2")$Study_id_f) # 12
n_distinct(subset(VEmetareg_data, clean_vaccine_dose=="Abdala (CIBG) x 3")$Study_id_f) # 1

VEmetareg_data <- VEmetareg_data %>%
  filter(
    clean_vaccine == "BioNTech/Pfizer BNT162b2" |
    clean_vaccine_dose == "Janssen Ad26.COV2.S x 1" |
    clean_vaccine == "Moderna mRNA-1273" |
    clean_vaccine == "Oxford/AstraZeneca ChAdOx1-S" |
    clean_vaccine == "Sinovac CoronaVac"
    )

nrow(VEmetareg_data) #423
n_distinct(VEmetareg_data$Study_id_f) #72

## Step 7: exclude variants with <10 unique studies
table(VEmetareg_data$variant_clean)
n_distinct(subset(VEmetareg_data, variant_clean=="alpha")$Study_id_f) # 15
n_distinct(subset(VEmetareg_data, variant_clean=="beta")$Study_id_f) # 4
n_distinct(subset(VEmetareg_data, variant_clean=="delta")$Study_id_f) # 40
n_distinct(subset(VEmetareg_data, variant_clean=="gamma")$Study_id_f) # 6
n_distinct(subset(VEmetareg_data, variant_clean=="mu")$Study_id_f) # 4
n_distinct(subset(VEmetareg_data, variant_clean=="omicron")$Study_id_f) # 31
n_distinct(subset(VEmetareg_data, variant_clean=="original")$Study_id_f) # 1

VEmetareg_data <- VEmetareg_data %>%
  filter(
    variant_clean == "alpha" |
    variant_clean == "delta" |
    variant_clean == "omicron"
    )
nrow(VEmetareg_data) #367
n_distinct(VEmetareg_data$Study_id_f) #65

# sort 
VEmetareg_data <- VEmetareg_data %>% arrange(outcome_final, variant_clean, clean_vaccine, Study_id_f)

# Save filtered VE database
write.csv(VEmetareg_data, "Data/Processed/COVID19VE_ineq_merged_filtered.csv")
```

Four vaccines (BNT162b2, Ad26.COV2.S, mRNA-1273, ChAdOx1-S, and  CoronaVac) and 3 variants (alpha, delta, omicron) met initial eligibility threshold for meta-regression


```{r}
# vaccine study distribution
table_vacc <- VEmetareg_data %>%
  distinct(Study_id_f, clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(clean_vaccine, total_study) %>%
  mutate(n = n(),
         `n(%)` = paste0(round(n / total_study * 100, 2), "%" , " (", n, "/",total_study, ")" )) %>% 
  ungroup() %>% 
  select(clean_vaccine, `n(%)`) %>% 
  distinct()

#variant of interest
table_VoI1 <- VEmetareg_data %>%
  distinct(Study_id_f, variant_clean) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(variant_clean, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0(round(n / total_study * 100, 2), "%" , " (", n, "/",total_study, ")" )) %>% 
  ungroup() %>% 
  select(variant_clean, `n(%)`) %>% 
  distinct()

table_VoI2 <- VEmetareg_data %>%
  distinct(Study_id_f, clean_vaccine, variant_clean) %>% 
  group_by(clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(clean_vaccine, variant_clean,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0(round(n / total_study * 100, 2), "%" , " (", n, "/",total_study, ")" )) %>% 
  ungroup() %>% 
  select(clean_vaccine, variant_clean, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = clean_vaccine, values_from = `n(%)`)

  
combined_VoI <- left_join(table_VoI1, table_VoI2)

#outcome
table_outcome1 <- VEmetareg_data %>%
  distinct(Study_id_f, outcome_final) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(outcome_final, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0(round(n / total_study * 100, 2), "%" , " (", n, "/",total_study, ")" )) %>% 
  ungroup() %>% 
  select(outcome_final, `n(%)`) %>% 
  distinct()

table_outcome2 <- VEmetareg_data %>%
  distinct(Study_id_f, clean_vaccine, outcome_final) %>% 
  group_by(clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(clean_vaccine, outcome_final,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0(round(n / total_study * 100, 2), "%" , " (", n, "/",total_study, ")" )) %>% 
  ungroup() %>% 
  select(clean_vaccine, outcome_final, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = clean_vaccine, values_from = `n(%)`)

  
combined_outcome <- left_join(table_outcome1, table_outcome2)


#GNI
table_GNI1 <- VEmetareg_data %>%
  distinct(Study_id_f, q_GNI_per_c_WB) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(q_GNI_per_c_WB, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0(round(n / total_study * 100, 2), "%" , " (", n, "/",total_study, ")" )) %>% 
  ungroup() %>% 
  select(q_GNI_per_c_WB, `n(%)`) %>% 
  distinct()

table_GNI2 <- VEmetareg_data %>%
  distinct(Study_id_f, clean_vaccine, q_GNI_per_c_WB) %>% 
  group_by(clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(clean_vaccine, q_GNI_per_c_WB,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0(round(n / total_study * 100, 2), "%" , " (", n, "/",total_study, ")" )) %>% 
  ungroup() %>% 
  select(clean_vaccine, q_GNI_per_c_WB, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = clean_vaccine, values_from = `n(%)`)

  
combined_GNI <- left_join(table_GNI1, table_GNI2)

```


#### Model-specific eligibility checks
```{r}
model_check <- VEmetareg_data %>%
    group_by(outcome_final, variant_clean, clean_vaccine) %>%
    filter (! duplicated(Study_id_f)) %>%
    summarise(
      n_GNI_Q1 = sum(q_GNI_per_c_WB==1),
      n_GNI_Q2 = sum(q_GNI_per_c_WB==2),
      n_GNI_Q3 = sum(q_GNI_per_c_WB==3),
      n_GNI_Q4 = sum(q_GNI_per_c_WB==4),
      sum = n(),
      sum_nonQ4 = sum(q_GNI_per_c_WB!=4)
    )
model_check$sufficient_n = model_check$sum>=10
model_check$n_nonQ4_over_2 = model_check$sum_nonQ4>=2
model_check$eligible = model_check$sufficient_n==TRUE & model_check$n_nonQ4_over_2==TRUE
write.csv(model_check, "Data/Processed/VE_model_check.csv")

# criteria met for 6 models in total
# severe - delta/omicron - BNT162b2, mRNA-1273, and ChAdOx1-S 

# apply filters
n_distinct(subset(VEmetareg_data, clean_vaccine =="Sinovac CoronaVac")$Study_id_f) # 7
n_distinct(subset(VEmetareg_data, variant_clean =="alpha")$Study_id_f) # 15
n_distinct(subset(VEmetareg_data, outcome_final !="severe")$Study_id_f) # 19
n_distinct(subset(VEmetareg_data, clean_vaccine == "Janssen Ad26.COV2.S")$Study_id_f) #10

VEmetareg_data <- VEmetareg_data %>%
  filter(
    clean_vaccine != "Sinovac CoronaVac" &
    outcome_final == "severe" &
    variant_clean != "alpha" &
    clean_vaccine != "Janssen Ad26.COV2.S"
    )
nrow(VEmetareg_data) #206
n_distinct(VEmetareg_data$Study_id_f) #51

# create max_duration_follow_up_days 
VEmetareg_data$max_follow_up_days = str_replace_all(VEmetareg_data$`max duration follow-up`, " weeks","")
VEmetareg_data$max_follow_up_days = str_replace_all(VEmetareg_data$max_follow_up_days, "~","")
VEmetareg_data$max_follow_up_days = as.numeric(VEmetareg_data$max_follow_up_days)*7

#Create variable number of subgroup by study for identifying studies with 2 or more subgroups
 VEmetareg_data <- VEmetareg_data %>%
    group_by(outcome_final, variant_clean, clean_vaccine, Study_id_f) %>% 
  mutate(
    n_estimates = n()
  ) %>% 
  ungroup()

# Save filtered VE database
write.csv(VEmetareg_data, "Data/Processed/COVID19VE_ineq_merged_filtered_eligible.csv")

# Manual addition of 5 variables:
#min_period_days - derived from timing_of_dose_days
#max_period_days - derived from timing_of_dose_days and `max duration follow-up`
#include - selection for studies with multiple estimates
#justification - reasons for selections
#sensitivity - binary flag for subpopulations >=60
```


```{r}
# sensitivity analysis
sensitivity_analysis <- read_csv("Data/Processed/COVID19VE_ineq_merged_filtered_eligible_selected.csv") %>% 
  filter(sensitivity=="1" & include=="yes")

sensitivity_model_check <- sensitivity_analysis %>%
    group_by(outcome_final, variant_clean, clean_vaccine) %>%
    filter (! duplicated(Study_id_f)) %>%
    summarise(
      n_GNI_Q1 = sum(q_GNI_per_c_WB==1),
      n_GNI_Q2 = sum(q_GNI_per_c_WB==2),
      n_GNI_Q3 = sum(q_GNI_per_c_WB==3),
      n_GNI_Q4 = sum(q_GNI_per_c_WB==4),
      sum = n(),
      sum_nonQ4 = sum(q_GNI_per_c_WB!=4)
    )
sensitivity_model_check$sufficient_n = sensitivity_model_check$sum>=10
sensitivity_model_check$n_nonQ4_over_2 = sensitivity_model_check$sum_nonQ4>=2
sensitivity_model_check$eligible = sensitivity_model_check$sufficient_n==TRUE & sensitivity_model_check$n_nonQ4_over_2==TRUE

write.csv(sensitivity_model_check, "Data/Processed/VE_sensitivity_model_check.csv")

# criteria met for 0 models 
```



```{r}
#Supplementary table 4
metareg_selected <- read_csv("Data/Processed/COVID19VE_ineq_merged_filtered_eligible_selected.csv") %>% 
  filter(include=="yes")

# vaccine study distribution
table_vacc <- metareg_selected %>%
  distinct(Study_id_f, clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(clean_vaccine, total_study) %>%
  mutate(n = n(),
         `n(%)` = paste0(n," (",round(n / total_study * 100, 0), "%)")) %>% 
  ungroup() %>% 
  select(clean_vaccine, `n(%)`) %>% 
  distinct()
table_vacc

#variant of interest
table_VoI1 <- metareg_selected %>%
  distinct(Study_id_f, variant_clean) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(variant_clean, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0(n," (",round(n / total_study * 100, 0), "%)")) %>% 
  ungroup() %>% 
  select(variant_clean, `n(%)`) %>% 
  distinct()

table_VoI2 <- metareg_selected %>%
  distinct(Study_id_f, clean_vaccine, variant_clean) %>% 
  group_by(clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(clean_vaccine, variant_clean,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0(n," (",round(n / total_study * 100, 0), "%)")) %>% 
  ungroup() %>% 
  select(clean_vaccine, variant_clean, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = clean_vaccine, values_from = `n(%)`)

  
combined_VoI <- left_join(table_VoI1, table_VoI2)
combined_VoI

#outcome
table_outcome1 <- metareg_selected %>%
  distinct(Study_id_f, outcome_final) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(outcome_final, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0(n," (",round(n / total_study * 100, 0), "%)")) %>% 
  ungroup() %>% 
  select(outcome_final, `n(%)`) %>% 
  distinct()

table_outcome2 <- metareg_selected %>%
  distinct(Study_id_f, clean_vaccine, outcome_final) %>% 
  group_by(clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(clean_vaccine, outcome_final,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0(n," (",round(n / total_study * 100, 0), "%)")) %>% 
  ungroup() %>% 
  select(clean_vaccine, outcome_final, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = clean_vaccine, values_from = `n(%)`)

  
combined_outcome <- left_join(table_outcome1, table_outcome2)
combined_outcome

#GNI
table_GNI1 <- metareg_selected %>%
  distinct(Study_id_f, q_GNI_per_c_WB) %>% 
  mutate(
    total_study=n_distinct(Study_id_f) 
  ) %>% 
  group_by(q_GNI_per_c_WB, total_study) %>% 
   mutate(n = n(),
         `n(%)` = paste0(n," (",round(n / total_study * 100, 0), "%)")) %>% 
  ungroup() %>% 
  select(q_GNI_per_c_WB, `n(%)`) %>% 
  distinct()

table_GNI2 <- metareg_selected %>%
  distinct(Study_id_f, clean_vaccine, q_GNI_per_c_WB) %>% 
  group_by(clean_vaccine) %>% 
  mutate(
    total_study=n_distinct(Study_id_f)) %>% 
  ungroup() %>% 
  group_by(clean_vaccine, q_GNI_per_c_WB,total_study) %>% 
 mutate(n = n(),
         `n(%)` = paste0(n," (",round(n / total_study * 100, 0), "%)")) %>% 
  ungroup() %>% 
  select(clean_vaccine, q_GNI_per_c_WB, `n(%)`) %>% 
  distinct() %>% 
  pivot_wider(names_from = clean_vaccine, values_from = `n(%)`)

  
combined_GNI <- left_join(table_GNI1, table_GNI2)
combined_GNI
```

