---
title: "7 Meta-regression results"
author: "Martina Pesce & Edward P. K. Parker"
date: "2023-10-12"
output: html_document
editor_options: 
  chunk_output_type: console
---

#### Inputs
- Processed/COVID19VE_ineq_merged_filtered_eligible_selected_metareg_var.xlsx: Processed meta-regression data

#### Outputs
- Meta-regression model results and plots for selected vaccine


#Moderna mRNA-1273

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# select vaccine
selected_vaccine = "Moderna mRNA-1273" # values: "BioNTech/Pfizer BNT162b2" "Oxford/AstraZeneca ChAdOx1-S" "Moderna mRNA-1273"
simple_name = "mRNA-1273" # for file outputs 

```



```{r}
library(tidyverse)
library(metafor)
library(dmetar)
library(forestplot)
library(patchwork)

VEmetareg_data_clean <- read_csv("Data/Processed/COVID19VE_ineq_merged_filtered_eligible_selected_metareg_var.csv") %>% 
  filter(include=="yes") %>% 
  mutate(publication_year = paste0( substring(`publication date`, nchar(`publication date`) - 3))) %>% 
  mutate(
    paper= paste0(
  `first author`, " (",
  publication_year, ")"
)
  ) %>% 
  arrange(country_clean)
  
n_distinct(VEmetareg_data_clean$Study_id_f)

# Update country names for plotting
VEmetareg_data_clean$country_clean[VEmetareg_data_clean$country_clean=="Hong Kong"] = "Hong Kong SAR"

# Simplify names for plotting
VEmetareg_data_clean$population_category[VEmetareg_data_clean$population_category=="immunocompetent adults"] = "immunocomp. adults"

# check values for vaccines, outcome, variant and quartile
table(VEmetareg_data_clean$clean_vaccine)
table(VEmetareg_data_clean$outcome_final) # all severe
table(VEmetareg_data_clean$variant_clean) # delta or omicron
table(VEmetareg_data_clean$q_GNI_per_c_WB) # 4, 3, 2

studies_countries_models <- VEmetareg_data_clean %>%
  group_by(variant_clean, clean_vaccine, q_GNI_per_c_WB) %>%
  summarise(
    countries= n_distinct(Country.Code),
    studies= n_distinct(Study_id_f),
    subgroup= n(),
    median_min_period = round(median(as.numeric(min_period_days), na.rm = TRUE)/7, 0),
    q25_min_period = round(quantile(as.numeric(min_period_days), 0.25, na.rm = TRUE)/7, 0),
    q75_min_period = round(quantile(as.numeric(min_period_days), 0.75, na.rm = TRUE)/7, 0),
    median_max_period = round(median(as.numeric(max_period_days), na.rm = TRUE)/7, 0),
    q25_max_period = round(quantile(as.numeric(max_period_days), 0.25, na.rm = TRUE)/7, 0),
    q75_max_period = round(quantile(as.numeric(max_period_days), 0.75, na.rm = TRUE)/7, 0)
  ) %>%
  mutate(
    median_min = paste0(median_min_period, " (", q25_min_period, "-", q75_min_period, ")"),
    median_max = paste0(median_max_period, " (", q25_max_period, "-", q75_max_period, ")")
  ) %>%
  select(-median_min_period, -q25_min_period, -q75_min_period, -median_max_period, -q25_max_period, -q75_max_period)
write.csv(studies_countries_models, "Data/Processed/meta_summary_counts.csv")
```

## MODEL 1.A: delta, severe 


```{r}
#Filter: delta, severe 

delta_severe <- VEmetareg_data_clean %>%
  filter (clean_vaccine == selected_vaccine & variant_clean=="delta") %>% 
  mutate(
    subgroup = as.factor(1:nrow(.)),
    Study_id_f = as.factor(Study_id_f),
    q_GNI_per_c_WB = factor(q_GNI_per_c_WB, levels=c("4", "3", "2")),
  )


```


# MODEL 1.B:omicron, severe 


```{r}

#Filter:omicron, severe 

omicron_severe <- VEmetareg_data_clean %>%
  filter (clean_vaccine == selected_vaccine & variant_clean=="omicron") %>% 
  mutate(
    subgroup = as.factor(1:nrow(.)),
    Study_id_f = as.factor(Study_id_f),
    q_GNI_per_c_WB = factor(q_GNI_per_c_WB, levels=c("4", "3", "2")),
    country_clean =as.factor(country_clean)
  )

```


#GNI

## Table 2. 

Primary COVID-19 vaccine series’ early effectiveness against severe COVID-19 disease by the variant of interest context and inequalities between countries (GNI per capita). RR= GNI per capita quartile pooled Risk Ratio. 95% CI= 95% Confidence interval: statistical range within which a population parameter is expected to lie. 95% PI= 95% predictive interval: estimate of the range within which the effects of future studies are expected to fall based on the existing evidence. I²= proportion of effect variability not caused by sampling error. 𝜏²= Variance components. Level 2(w)= heterogeneity within countries. Level 3(b)=  heterogeneity between countries.


```{r}
### MODEL delta_severe: GNI
delta_severe.GNI <-
  rma.mv(
    yi = ln.RR,
    V = (SE.ln.RR)^2,
    slab = `first author`,
    data = delta_severe, 
    mods = ~ q_GNI_per_c_WB, 
    random = ~ 1 | iso3c /  subgroup,
    test = "t",
    method = "REML"
  )


#Estimate, CI, PI (RR)
pred.delta_severe.GNI <- as.data.frame(predict(delta_severe.GNI, addx=TRUE, transf=exp) ) 
pred.delta_severe.GNI <-  distinct(pred.delta_severe.GNI)
pred.delta_severe.GNI$GNI = 4
pred.delta_severe.GNI$GNI[pred.delta_severe.GNI$X.q_GNI_per_c_WB3==1] = 3
pred.delta_severe.GNI$GNI[pred.delta_severe.GNI$X.q_GNI_per_c_WB2==1] = 2
pred.delta_severe.GNI$VoI <- rep("delta", nrow(pred.delta_severe.GNI))
# pred.delta_severe.GNI$subgroup <- 
# pred.delta_severe.GNI$country <- 


#i^2
I2_lvl2.delta_severe.GNI<- mlm.variance.distribution(delta_severe.GNI)$results[2,2] #lvl 2 i^2
I2_lvl3.delta_severe.GNI<- mlm.variance.distribution(delta_severe.GNI)$results[3,2] #lvl 3 i^2

o2_lvl2.delta_severe.GNI <- sqrt(delta_severe.GNI$sigma2[2]) #within cluster sigma (level 2)
o2_lvl3.delta_severe.GNI <- sqrt(delta_severe.GNI$sigma2[1]) #between cluster sigma (level 3)

#moderator p-value
QMp_value.delta_severe.GNI <- delta_severe.GNI$QMp
#m1.GNI$k.w

I_o_p_delta_severe.GNI <- data.frame(
  QMp_value = QMp_value.delta_severe.GNI, 
  I2_lvl2 = I2_lvl2.delta_severe.GNI,
   I2_lvl3 = I2_lvl3.delta_severe.GNI,
   o2_lvl2 = o2_lvl2.delta_severe.GNI,
   o2_lvl3 = o2_lvl3.delta_severe.GNI
 )

Sum_delta_severe.GNI <- cbind.data.frame(pred.delta_severe.GNI, I_o_p_delta_severe.GNI) %>%
  select(-starts_with("X.")) %>% 
  mutate(RR_summary = 
           paste0(
             format(round(pred,2),nsmall=2)," (",
             format(round(ci.lb,2),nsmall=2),"-",
             format(round(ci.ub,2),nsmall=2),")"
             ),
         VE_summary = 
           paste0(
             format(round((1-pred)*100,0),nsmall=0)," (",
             format(round((1-ci.ub)*100,0),nsmall=0),"-",
             format(round((1-ci.lb)*100,0),nsmall=0),")"
             )
         ) 
Sum_delta_severe.GNI # order is GNI2 - row 1, GNI 3 - row 2, GNI 4 - row 1

# summary metrics for GNI quartile 2
delta_rr_GNI_2 = Sum_delta_severe.GNI$pred[3]
delta_lci_GNI_2 = Sum_delta_severe.GNI$ci.lb[3]
delta_uci_GNI_2 = Sum_delta_severe.GNI$ci.ub[3]
delta_RR_summary_GNI_2 = Sum_delta_severe.GNI$RR_summary[3]

# summary metrics for GNI quartile 3
delta_rr_GNI_3 = Sum_delta_severe.GNI$pred[2]
delta_lci_GNI_3 = Sum_delta_severe.GNI$ci.lb[2]
delta_uci_GNI_3 = Sum_delta_severe.GNI$ci.ub[2]
delta_RR_summary_GNI_3 = Sum_delta_severe.GNI$RR_summary[2]

# summary metrics for GNI quartile 4
delta_rr_GNI_4 = Sum_delta_severe.GNI$pred[1]
delta_lci_GNI_4 = Sum_delta_severe.GNI$ci.lb[1]
delta_uci_GNI_4 = Sum_delta_severe.GNI$ci.ub[1]
delta_RR_summary_GNI_4 = Sum_delta_severe.GNI$RR_summary[1]
```




```{r}
### MODEL omicron_severe: GNI
omicron_severe.GNI <-
  rma.mv(
    yi = ln.RR,
    V = (SE.ln.RR)^2,
    slab = `first author`,
    data = omicron_severe, 
    mods = ~ q_GNI_per_c_WB, 
    random = ~ 1 | iso3c /  subgroup,
    test = "t",
    method = "REML"
  )


#Estimate, CI, PI (RR)
pred.omicron_severe.GNI <- as.data.frame(predict(omicron_severe.GNI, addx=TRUE, transf=exp) ) 
pred.omicron_severe.GNI <-  distinct(pred.omicron_severe.GNI)
pred.omicron_severe.GNI$GNI = 4
pred.omicron_severe.GNI$GNI[pred.omicron_severe.GNI$X.q_GNI_per_c_WB3==1] = 3
pred.omicron_severe.GNI$GNI[pred.omicron_severe.GNI$X.q_GNI_per_c_WB2==1] = 2
pred.omicron_severe.GNI$VoI <- rep("omicron", nrow(pred.omicron_severe.GNI))



#i^2
I2_lvl2.omicron_severe.GNI<- mlm.variance.distribution(omicron_severe.GNI)$results[2,2] #lvl 2 i^2
I2_lvl3.omicron_severe.GNI<- mlm.variance.distribution(omicron_severe.GNI)$results[3,2] #lvl 3 i^2

o2_lvl2.omicron_severe.GNI <- sqrt(omicron_severe.GNI$sigma2[2]) #within cluster sigma (level 2)
o2_lvl3.omicron_severe.GNI <- sqrt(omicron_severe.GNI$sigma2[1]) #between cluster sigma (level 3)

#moderator p-value
QMp_value.omicron_severe.GNI <- omicron_severe.GNI$QMp
#m1.GNI$k.w

I_o_p_omicron_severe.GNI <- data.frame(
  QMp_value = QMp_value.omicron_severe.GNI, 
  I2_lvl2 = I2_lvl2.omicron_severe.GNI,
   I2_lvl3 = I2_lvl3.omicron_severe.GNI,
   o2_lvl2 = o2_lvl2.omicron_severe.GNI,
   o2_lvl3 = o2_lvl3.omicron_severe.GNI
 )

Sum_omicron_severe.GNI <- cbind.data.frame(pred.omicron_severe.GNI, I_o_p_omicron_severe.GNI) %>%
   select(-starts_with("X.")) %>% 
    mutate(RR_summary = 
           paste0(
             format(round(pred,2),nsmall=2)," (",
             format(round(ci.lb,2),nsmall=2),"-",
             format(round(ci.ub,2),nsmall=2),")"
             ),
         VE_summary = 
           paste0(
             format(round((1-pred)*100,0),nsmall=0)," (",
             format(round((1-ci.ub)*100,0),nsmall=0),"-",
             format(round((1-ci.lb)*100,0),nsmall=0),")"
             )
         ) 
Sum_omicron_severe.GNI # order is GNI2 - row 3, GNI 3 - row 1, GNI 4 - row 2

# summary metrics for GNI quartile 2
omicron_rr_GNI_2 = Sum_omicron_severe.GNI$pred[3]
omicron_lci_GNI_2 = Sum_omicron_severe.GNI$ci.lb[3]
omicron_uci_GNI_2 = Sum_omicron_severe.GNI$ci.ub[3]
omicron_RR_summary_GNI_2 = Sum_omicron_severe.GNI$RR_summary[3]

# summary metrics for GNI quartile 3
omicron_rr_GNI_3 = Sum_omicron_severe.GNI$pred[1]
omicron_lci_GNI_3 = Sum_omicron_severe.GNI$ci.lb[1]
omicron_uci_GNI_3 = Sum_omicron_severe.GNI$ci.ub[1]
omicron_RR_summary_GNI_3 = Sum_omicron_severe.GNI$RR_summary[1]

# summary metrics for GNI quartile 4
omicron_rr_GNI_4 = Sum_omicron_severe.GNI$pred[2]
omicron_lci_GNI_4 = Sum_omicron_severe.GNI$ci.lb[2]
omicron_uci_GNI_4 = Sum_omicron_severe.GNI$ci.ub[2]
omicron_RR_summary_GNI_4 = Sum_omicron_severe.GNI$RR_summary[2]

```

```{r}
#Table severe (Omicron + Delta)

#Data meta GNI
meta_GNI <- rbind(Sum_omicron_severe.GNI, Sum_delta_severe.GNI )

meta_GNI$vaccine = selected_vaccine
meta_GNI$clean_vaccine = simple_name
write.csv(meta_GNI, paste0("Data/Processed/meta_summary_",simple_name,".csv"))
```

## Forest plots

#### Delta 
```{r}
# Code country as character
delta_severe$country_clean = as.character(delta_severe$country_clean)

# Created collated RR (95% CI) values for table
delta_severe$RR_full =
  paste0(
    format(round(delta_severe$RR,2),nsmall=2)," (",
    format(round(delta_severe$RR.LCL,2),nsmall=2),"-",
    format(round(delta_severe$RR.UCL,2),nsmall=2),")"
    )

# Create subgroups for each quartile
delta_gni_2 = subset(delta_severe, q_GNI_per_c_WB==2)
delta_gni_3 = subset(delta_severe, q_GNI_per_c_WB==3)
delta_gni_4 = subset(delta_severe, q_GNI_per_c_WB==4)

# Class numeric
delta_gni_2$max_period_days <- as.numeric(delta_gni_2$max_period_days)
delta_gni_3$max_period_days <- as.numeric(delta_gni_3$max_period_days)
delta_gni_4$max_period_days <- as.numeric(delta_gni_4$max_period_days)

# construct table
titles = c("Study",
           "GNI quartile 2",delta_gni_2$paper,"Summary",NA,
           "GNI quartile 3",delta_gni_3$paper,"Summary",NA,
           "GNI quartile 4",delta_gni_4$paper,"Summary")
country = c("Country",NA,
            delta_gni_2$country_clean,NA,NA,NA,
            delta_gni_3$country_clean,NA,NA,NA,
            delta_gni_4$country_clean,NA)
population = c("Population",NA,
               delta_gni_2$population_category,NA,NA,NA,
               delta_gni_3$population_category,NA,NA,NA,
               delta_gni_4$population_category,NA)
follow_up = c("Max. follow-up (w)",NA,
              round(delta_gni_2$max_period_days/7,0),NA,NA,NA,
              round(delta_gni_3$max_period_days/7,0),NA,NA,NA,
              round(delta_gni_4$max_period_days/7,0),NA)
RRs = c("RR (95% CI)",NA,
        delta_gni_2$RR_full,delta_RR_summary_GNI_2,NA,NA,
        delta_gni_3$RR_full,delta_RR_summary_GNI_3,NA,NA,
        delta_gni_4$RR_full,delta_RR_summary_GNI_4)
tabletext = cbind(titles, country, population, follow_up, RRs)


# list RRs
m = c(NA,NA,
      delta_gni_2$RR,delta_rr_GNI_2,NA,NA,
      delta_gni_3$RR,delta_rr_GNI_3,NA,NA,
      delta_gni_4$RR,delta_rr_GNI_4)
l = c(NA,NA,
      delta_gni_2$RR.LCL,delta_lci_GNI_2,NA,NA,
      delta_gni_3$RR.LCL,delta_lci_GNI_3,NA,NA,
      delta_gni_4$RR.LCL,delta_lci_GNI_4)
u = c(NA,NA,
      delta_gni_2$RR.UCL,delta_uci_GNI_2,NA,NA,
      delta_gni_3$RR.UCL,delta_uci_GNI_3,NA,NA,
      delta_gni_4$RR.UCL,delta_uci_GNI_4)


# truncate variables for plotting
m[m<0.01] = 0.01
l[l<0.01] = 0.01
u[u>1] = 1
u[is.na(u)] = 1

# generate forest plot
forestplot(tabletext,m,l,u,
    is.summary=c(TRUE,TRUE,!is.na(tabletext[3:nrow(tabletext),1]) & 
                   (tabletext[3:nrow(tabletext),1]=="Summary" | 
                      tabletext[3:nrow(tabletext),1]=="GNI quartile 3" |
                      tabletext[3:nrow(tabletext),1]=="GNI quartile 4")),
    zero=NA,
    xlog=TRUE,
    xticks=c(0.01,0.1,0.2,0.5,1),
    colgap=unit(5,"mm"),
    txt_gp = fpTxtGp(ticks = gpar(cex=1), xlab = gpar(cex=1)),
    xlab="RR"
    )

```

#### Omicron
```{r}
# Code country as character
omicron_severe$country_clean = as.character(omicron_severe$country_clean)

# Created collated RR (95% CI) values for table
omicron_severe$RR_full =
  paste0(
    format(round(omicron_severe$RR,2),nsmall=2)," (",
    format(round(omicron_severe$RR.LCL,2),nsmall=2),"-",
    format(round(omicron_severe$RR.UCL,2),nsmall=2),")"
    )

# Create subgroups for each quartile
omicron_gni_2 = subset(omicron_severe, q_GNI_per_c_WB==2)
omicron_gni_3 = subset(omicron_severe, q_GNI_per_c_WB==3)
omicron_gni_4 = subset(omicron_severe, q_GNI_per_c_WB==4)

# Class numeric
omicron_gni_2$max_period_days <- as.numeric(omicron_gni_2$max_period_days)
omicron_gni_3$max_period_days <- as.numeric(omicron_gni_3$max_period_days)
omicron_gni_4$max_period_days <- as.numeric(omicron_gni_4$max_period_days)

# construct table
titles = c("Study",
           "GNI quartile 2",omicron_gni_2$paper,"Summary",NA,
           "GNI quartile 3",omicron_gni_3$paper,"Summary",NA,
           "GNI quartile 4",omicron_gni_4$paper,"Summary")
country = c("Country",NA,
            omicron_gni_2$country_clean,NA,NA,NA,
            omicron_gni_3$country_clean,NA,NA,NA,
            omicron_gni_4$country_clean,NA)
population = c("Population",NA,
               omicron_gni_2$population_category,NA,NA,NA,
               omicron_gni_3$population_category,NA,NA,NA,
               omicron_gni_4$population_category,NA)
follow_up = c("Max. follow-up (w)",NA,
              round(omicron_gni_2$max_period_days/7,0),NA,NA,NA,
              round(omicron_gni_3$max_period_days/7,0),NA,NA,NA,
              round(omicron_gni_4$max_period_days/7,0),NA)
RRs = c("RR (95% CI)",NA,
        omicron_gni_2$RR_full,omicron_RR_summary_GNI_2,NA,NA,
        omicron_gni_3$RR_full,omicron_RR_summary_GNI_3,NA,NA,
        omicron_gni_4$RR_full,omicron_RR_summary_GNI_4)
tabletext = cbind(titles, country, population, follow_up, RRs)

# list RRs
m = c(NA,NA,
      omicron_gni_2$RR,omicron_rr_GNI_2,NA,NA,
      omicron_gni_3$RR,omicron_rr_GNI_3,NA,NA,
      omicron_gni_4$RR,omicron_rr_GNI_4)
l = c(NA,NA,
      omicron_gni_2$RR.LCL,omicron_lci_GNI_2,NA,NA,
      omicron_gni_3$RR.LCL,omicron_lci_GNI_3,NA,NA,
      omicron_gni_4$RR.LCL,omicron_lci_GNI_4)
u = c(NA,NA,
      omicron_gni_2$RR.UCL,omicron_uci_GNI_2,NA,NA,
      omicron_gni_3$RR.UCL,omicron_uci_GNI_3,NA,NA,
      omicron_gni_4$RR.UCL,omicron_uci_GNI_4)

# truncate variables for plotting
m[m<0.01] = 0.01
l[l<0.01] = 0.01
u[u>1] = 1
u[is.na(u)] = 1

# generate forest plot
forestplot(tabletext,m,l,u,
    is.summary=c(TRUE,TRUE,!is.na(tabletext[3:nrow(tabletext),1]) & 
                   (tabletext[3:nrow(tabletext),1]=="Summary" | 
                      tabletext[3:nrow(tabletext),1]=="GNI quartile 3" |
                      tabletext[3:nrow(tabletext),1]=="GNI quartile 4")),
    zero=NA,
    xlog=TRUE,
    xticks=c(0.01,0.1,0.2,0.5,1),
    colgap=unit(5,"mm"),
    txt_gp = fpTxtGp(ticks = gpar(cex=1), xlab = gpar(cex=1)),
    xlab="RR"
    )

```

